%CHAPTER 19
%\chapterfigure[scale=0.2]{Durer/Durer_Revelation_Four_Riders.jpg}{The Four Horsemen. Albrecht Dürer, 1498.}

\chapter{The Marriage of the Lamb}
\lettrine[lines=3,slope=0.5em, lraise=0.3]{\textcolor{red}{A}}{\ fter} these things I heard as it were a great voice of a great multitude%
	\cbiblefoot{Daniel}{10:5-6}{I lifted up mine eyes, and looked, and, behold, a man\ldots and the voice of his words like the voice of a multitude}
 in heaven, saying,
\zz \begin{verse}
Hallelujah;\\
Salvation, and glory, and power, belong to our God:\\*\vin
\vleftofline{\vnum{2}}for true and righteous are his judgments;\\
for he hath judged the great harlot,\\*\vin
her that corrupted the earth with her fornication,\\
and he hath avenged the blood of his servants at her hand.%
	\cbiblefoottrio{Deuteronomy}{32:43}{Rejoice, O ye nations, with his people: For he will avenge the blood of his servants, And will render vengeance to his adversaries, And will make expiation for his land, for his people}%
				{IIKings}{9:7}{And thou shalt smite the house of Ahab thy master, that I may avenge the blood of my servants the prophets, and the blood of all the servants of Jehovah, at the hand of Jezebel}%
				{Psalms}{79:10}{Wherefore should the nations say, Where is their God? Let the avenging of the blood of thy servants which is shed be known among the nations in our sight}
\end{verse}

\noindent
\vnum{3}And a second time they say,
\begin{verse}
Hallelujah.\\
And her smoke goeth up for ever and ever.%
	\cbiblefoot{Isaiah}{34:9-10}{And the streams of Edom shall be turned into pitch, and the dust thereof into brimstone, and the land thereof shall become burning pitch. It shall not be quenched night nor day; the smoke thereof shall go up for ever; from generation to generation it shall lie waste; none shall pass through it for ever and ever}
\end{verse}
\vnum{4}And the four and twenty elders and the four living creatures fell down and worshipped God that sitteth on the throne, saying, 
\begin{verse}
Amen; Hallelujah. 
\end{verse}
\vnum{5}And a voice came forth from the throne, saying,
\begin{verse}
Give praise to our God,\\*\vin
all ye his servants\\*
ye that fear him,\\*\vin
the small and the great.%
	\footnote{\cbibleref{Psalms}{115:13}{He will bless them that fear Jehovah, both small and great}}
\end{verse}
\vnum{6}And I heard as it were the voice of a great multitude, and as the voice of many waters, and as the voice of mighty thunders, saying,
\begin{verse}
Hallelujah:\\
for the Lord our God, the Almighty, reigneth.%
	\footnote{\cbibleref{Psalms}{93:1}{Jehovah reigneth; he is clothed with majesty; Jehovah is clothed with strength; he hath girded himself therewith: The world also is established, that it cannot be moved} %
			 \cbiblechvs{Psalms}{99:1}{Jehovah reigneth; let the peoples tremble: He sitteth above the cherubim; let the earth be moved} %
			 \cbibleref{IChronicles}{16:31}{Let the heavens be glad, and let the earth rejoice; And let them say among the nations, Jehovah reigneth} %
			 \cbibleref{Zechariah}{14:9}{And Jehovah shall be King over all the earth: in that day shall Jehovah be one, and his name one}}\\
\vleftofline{\vnum{7}}Let us rejoice and be exceeding glad,%
	\cbiblefoot{Psalms}{118:24}{This is the day which Jehovah hath made; We will rejoice and be glad in it}
 and let us give the glory unto him:\\
for the marriage of the Lamb is come, and his wife hath made herself ready.\\
\vleftofline{\vnum{8}}And it was given unto her that she should array herself in fine linen, bright and pure
\end{verse}
For the fine linen is the righteous acts of the saints.%
	\cbiblefoot{Isaiah}{61:10}{I will greatly rejoice in Jehovah, my soul shall be joyful in my God; for he hath clothed me with the garments of salvation, he hath covered me with the robe of righteousness, as a bridegroom decketh himself with a garland, and as a bride adorneth herself with her jewels}

\vnum{9}And he saith unto me, Write, Blessed are they that are bidden to the marriage supper of the Lamb. And he saith unto me, These are true words of God. 
\vnum{10}And I fell down before his feet to worship him. And he saith unto me, See thou do it not: I am a fellow-servant with thee and with thy brethren that hold the testimony of Jesus: worship God: for the testimony of Jesus is the spirit of prophecy.

\vnum{11}And I saw the heaven opened;%
	\cbiblefoot{Ezekiel}{1:1}{Now it came to pass\ldots that the heavens were opened, and I saw visions of God}
 and behold, a white horse, and he that sat thereon called Faithful and True; and in righteousness he doth judge%
	\footnote{\cbibleref{Psalms}{72:2}{He will judge thy people with righteousness, and thy poor with justice} %
			  \cbiblechvs{Psalms}{96:13}{Before Jehovah; for he cometh, for he cometh to judge the earth: he will judge the world with righteousness, and the peoples with his truth} %
			  \cbiblechvs{Psalms}{98:9}{Before Jehovah; for he cometh to judge the earth: he will judge the world with righteousness, and the peoples with equity} %
			  \cbibleref{Isaiah}{11:4-5}{But with righteousness shall he judge the poor, and decide with equity for the meek of the earth; and he shall smite the earth with the rod of his mouth; and with the breath of his lips shall he slay the wicked. And righteousness shall be the girdle of his waist, and faithfulness the girdle of his loins}}
 and make war. 
\vnum{12}And his eyes are a flame of fire,%
	\cbiblefoot{Daniel}{10:6}{His body also was like the beryl, and his face as the appearance of lightning, and his eyes as flaming torches, and his arms and his feet like unto burnished brass, and the voice of his words like the voice of a multitude}
 and upon his head are many diadems; and he hath a name written which no one knoweth but he himself.%
	\cbiblefoot{Isaiah}{62:2-3}{And the nations shall see thy righteousness, and all kings thy glory, and thou shalt be called by a new name, which the mouth of Jehovah shall name. Thou shalt also be a crown of beauty in the hand of Jehovah, and a royal diadem in the hand of thy God}
\vnum{13}And he is arrayed in a garment sprinkled with blood:%
	\cbiblefoot{Isaiah}{63:1-6}{Who is this that cometh from Edom, with dyed garments from Bozrah? this that is glorious in his apparel, marching in the greatness of his strength? I that speak in righteousness, mighty to save. Wherefore art thou red in thine apparel, and thy garments like him that treadeth in the winevat? I have trodden the winepress alone; and of the peoples there was no man with me: yea, I trod them in mine anger, and trampled them in my wrath; and their lifeblood is sprinkled upon my garments, and I have stained all my raiment. For the day of vengeance was in my heart, and the year of my redeemed is come. And I looked, and there was none to help; and I wondered that there was none to uphold: therefore mine own arm brought salvation unto me; and my wrath, it upheld me. And I trod down the peoples in mine anger, and made them drunk in my wrath, and I poured out their lifeblood on the earth}
 and his name is called The Word of God. 
\vnum{14}And the armies which are in heaven followed him upon white horses, clothed in fine linen, white and pure. 
\vnum{15}And out of his mouth proceedeth a sharp sword,%
	\cbiblefoot{Isaiah}{49:2}{And he hath made my mouth like a sharp sword; in the shadow of his hand hath he hid me: and he hath made me a polished shaft; in his quiver hath he kept me close}
 that with it he should smite the nations: and he shall rule them with a rod of iron:%
	\cbiblefoot{Psalms}{2:9}{Thou shalt break them with a rod of iron; thou shalt dash them in pieces like a potter’s vessel}
 and he treadeth the winepress of the fierceness of the wrath of God,%
	\footnote{cf. \cbiblerefonly{Isaiah}{63:1-6} quoted in v. 13 footnote}
 the Almighty. 
\vnum{16}And he hath on his garment and on his thigh a name written, King of Kings, and Lord of Lords.%
	\cbiblefoot{Deuteronomy}{10:17}{For Jehovah your God, he is God of gods, and Lord of lords, the great God, the mighty, and the terrible, who regardeth not persons, nor taketh reward}

\vnum{17}And I saw an angel standing in the sun; and he cried with a loud voice, saying to all the birds that fly in mid heaven,%
	\cbiblefootduosbch{Ezekiel}{39:4}{Thou shalt fall upon the mountains of Israel, thou, and all thy hordes, and the peoples that are with thee: I will give thee unto the ravenous birds of every sort, and to the beasts of the field to be devoured}%
							{39:17-20}{And thou, son of man, thus saith the Lord Jehovah: Speak unto the birds of every sort, and to every beast of the field, Assemble yourselves, and come; gather yourselves on every side to my sacrifice that I do sacrifice for you, even a great sacrifice upon the mountains of Israel, that ye may eat flesh and drink blood. Ye shall eat the flesh of the mighty, and drink the blood of the princes of the earth, of rams, of lambs, and of goats, of bullocks, all of them fatlings of Bashan. And ye shall eat fat till ye be full, and drink blood till ye be drunken, of my sacrifice which I have sacrificed for you. And ye shall be filled at my table with horses and chariots, with mighty men, and with all men of war, saith the Lord Jehovah}
 Come and be gathered together unto the great supper of God; 
\vnum{18}that ye may eat the flesh of kings, and the flesh of captains, and the flesh of mighty men, and the flesh of horses and of them that sit thereon, and the flesh of all men, both free and bond, and small and great.

\vnum{19}And I saw the beast, and the kings of the earth, and their armies, gathered together to make war against him that sat upon the horse, and against his army.%
	\cbiblefoot{Psalms}{2:2}{The kings of the earth set themselves, and the rulers take counsel together, against Jehovah, and against his anointed}
\vnum{20}And the beast was taken,%
	\cbiblefoot{Daniel}{7:11-12}{I beheld at that time because of the voice of the great words which the horn spake; I beheld even till the beast was slain, and its body destroyed, and it was given to be burned with fire. And as for the rest of the beasts, their dominion was taken away: yet their lives were prolonged for a season and a time}
 and with him the false prophet that wrought the signs in his sight, wherewith he deceived them that had received the mark of the beast and them that worshipped his image: they two were cast alive%
	\cbiblefootduo{Numbers}{16:33}{So they, and all that appertained to them, went down alive into Sheol: and the earth closed upon them, and they perished from among the assembly}%
				  {Psalms}{55:15}{Let death come suddenly upon them, let them go down alive into Sheol; for wickedness is in their dwelling, in the midst of them}
 into the lake of fire that burneth with brimstone:%
	\cbiblefoot{Isaiah}{30:33}{For a Topheth is prepared of old; yea, for the king it is made ready; he hath made it deep and large; the pile thereof is fire and much wood; the breath of Jehovah, like a stream of brimstone, doth kindle it}
\vnum{21}and the rest were killed with the sword of him that sat upon the horse, even the sword which came forth out of his mouth:%
	\cbiblefoot{Isaiah}{11:4}{But with righteousness shall he judge the poor, and decide with equity for the meek of the earth; and he shall smite the earth with the rod of his mouth; and with the breath of his lips shall he slay the wicked}
 and all the birds were filled with their flesh.%
	\footnote{cf. \cbiblerefonly{Ezekiel}{39:4, 17-20} quoted in v. 17 footnote}
