\documentclass[11pt,a4paper,twocolumn,twoside]{book}

\usepackage{ebgaramond}
\usepackage[T1]{fontenc}
\usepackage{titlesec}
\usepackage[dvipsnames]{xcolor}
\usepackage{lettrine}
\usepackage[english]{babel}
%\usepackage[columns=3]{marginfootnotes}

% lettrine drop caps
\renewcommand{\LettrineFontHook}{\color[gray]{0.5}}

%reference quotes
\newcommand{\bibref}[2]{{#1}\ldots\textit{{#2}\ldots}}
\newcommand{\bibfootnote}[2]{\footnote{\bibref{#1}{#2}}}
\newcommand{\dibibref}[4]{\bibref{#1}{#2}\bibref{#3}{#4}}
\newcommand{\dibibfootnote}[4]{\footnote{\dibibref{#1}{#2}{#3}{#4}}}
\newcommand{\tribibref}[6]{\bibref{#1}{#2}\bibref{#3}{#4}\bibref{#5}{#6}}
\newcommand{\tribibfootnote}[6]{\footnote{\tribibref{#1}{#2}{#3}{#4}{#5}{#6}}}

%verse numbers
\newcommand{\vn}[1]{{\scriptsize\textbf{#1}}}

%page layout
%\addtolength{\textheight}{-5cm}
%\addtolength{\textwidth}{-8cm}
%\addtolength{\marginparwidth}{2cm}
%\addtolength{\hoffset}{1cm}
%\addtolength{\evensidemargin}{1cm}
%\addtolength{\oddsidemargin}{2cm}

\begin{document}
\frontmatter
\lettrine[lines=4,lraise=0.1]{T}{he Revelation} of Jesus Christ, which God gave unto him, to shew unto his servants things which must shortly come to pass; and he sent and signified it by his angel unto his servant John:

\vn{2} Who bare record of the word of God, and of the testimony of Jesus Christ, and of all things that he saw.

\vn{3} Blessed is he that readeth, and they that hear the words of this prophecy, and keep those things which are written therein: for the time is at hand.

\vn{4} John to the seven churches which are in Asia: Grace be unto you, and peace, from him which is, and which was, and which is to come; and from the seven Spirits which are before his throne;

\vn{5} And from Jesus Christ, who is the faithful witness, and the first begotten of the dead, and the prince of the kings of the earth. Unto him that loved us, and washed us from our sins in his own blood,

\vn{6} And hath made us kings and priests unto God and his Father; to him be glory and dominion for ever and ever. Amen.\bibfootnote{Ex 19.6}{And ye shall be unto me a kingdom of priests, and an holy nation}	

\vn{7} Behold, he cometh with clouds\bibfootnote{Dan 7.13}{one like the Son of man came with the clouds of heaven}; and every eye shall see him\bibfootnote{Is 40.5}{the glory of the LORD shall be revealed, And all flesh shall see it together}, and they also which pierced him: and all kindreds of the earth shall wail because of him\bibfootnote{Zech 12.10}{they shall look upon me whom they have pierced, And they shall mourn for him}. Even so, Amen. 

\vn{8} I am Alpha and Omega, the beginning and the ending\dibibfootnote{Is 41.4}{I the LORD, the first, And with the last; I am he}{Is 44.6}{I am the first, and I am the last; And beside me there is no God}, saith the Lord, which is, and which was, and which is to come, the Almighty.

\vn{9} I John, who also am your brother, and companion in tribulation, and in the kingdom and patience of Jesus Christ, was in the isle that is called Patmos, for the word of God, and for the testimony of Jesus Christ.

\vn{10} I was in the Spirit on the Lord’s day, and heard behind me a great voice, as of a trumpet,

\vn{11} Saying, I am Alpha and Omega, the first and the last: and, What thou seest, write in a book, and send it unto the seven churches which are in Asia; unto Ephesus, and unto Smyrna, and unto Pergamos, and unto Thyatira, and unto Sardis, and unto Philadelphia, and unto Laodicea.

\vn{12} And I turned to see the voice that spake with me. And being turned, I saw seven golden candlesticks\bibfootnote{Zech 4.2}{a candlestick all of gold, with a bowl upon the top of it, and his seven lamps thereon};

\vn{13} And in the midst of the seven candlesticks one like unto the Son of man, clothed with a garment down to the foot, and girt about the paps with a golden girdle\bibfootnote{Dan 10.5}{a certain man clothed in linen, whose loins were girded with fine gold}.

\vn{14} His head and his hairs were white like wool, as white as snow\bibfootnote{Dan 7.9-10}{the Ancient of days did sit, whose garment was white as snow, and the hair of his head like the pure wool: his throne was like the fiery flame, and his wheels as burning fire. A fiery stream issued and came forth from before him}; and his eyes were as a flame of fire\bibfootnote{Dan 10.6}{his eyes as lamps of fire};

\vn{15} And his feet like unto fine brass, as if they burned in a furnace\tribibfootnote{Dan 10.6}{his arms and his feet were like in colour to polished brass}{Eze 1.27}{I saw the as the colour of amber, as the appearance of fire round about within it, from the appearance of his loins even upward, and from the appearance of his loins even downward, I saw as it were the appearance of fire}{Eze 8.2}{a likeness as the appearance of fire: from the appearance of his loins even downward, fire; and from his loins even upward, as the appearance of brightness, as the colour of amber}; and his voice as the sound of many waters\bibfootnote{Eze 43.2}{his voice was like a noise of many waters}.

\vn{16} And he had in his right hand seven stars: and out of his mouth went a sharp twoedged sword\bibfootnote{Is 49.2}{he hath made my mouth like a sharp sword}: and his countenance was as the sun shineth in his strength.

\vn{17} And when I saw him, I fell at his feet as dead\dibibfootnote{Dan 8.17}{when he came, I was afraid, and fell upon my face}{Dan 10.8-10}{there remained no strength in me: for my comeliness was turned in me into corruption, and I retained no strength\ldots then was I in a deep sleep on my face, and my face toward the ground}. And he laid his right hand upon me, saying unto me\bibfootnote{Dan 10.9-10}{Yet I heard the voice of his words\ldots And, behold, an hand touched me, which set me upon my knees and upon the palms of my hands}, Fear not\bibfootnote{Dan 10.12}{Then he said unto me, Fear not, Daniel}; I am the first and the last\bibfootnote{Is 44.6}{I am the first, and I am the last; And besides me there is no God}:

\vn{18} I am he that liveth, and was dead; and, behold, I am alive for evermore, Amen; and have the keys of hell and of death.

\vn{19} Write the things which thou hast seen, and the things which are, and the things which shall be hereafter;

\vn{20} The mystery of the seven stars which thou sawest in my right hand, and the seven golden candlesticks. The seven stars are the angels of the seven churches: and the seven candlesticks which thou sawest are the seven churches.

Unto the angel of the church of Ephesus write; These things saith he that holdeth the seven stars in his right hand, who walketh in the midst of the seven golden candlesticks;

2 I know thy works, and thy labour, and thy patience, and how thou canst not bear them which are evil: and thou hast tried them which say they are apostles, and are not, and hast found them liars:

3 And hast borne, and hast patience, and for my name's sake hast laboured, and hast not fainted.

4 Nevertheless I have somewhat against thee, because thou hast left thy first love.

5 Remember therefore from whence thou art fallen, and repent, and do the first works; or else I will come unto thee quickly, and will remove thy candlestick out of his place, except thou repent.

6 But this thou hast, that thou hatest the deeds of the Nicolaitans, which I also hate.

7 He that hath an ear, let him hear what the Spirit saith unto the churches; To him that overcometh will I give to eat of the tree of life, which is in the midst of the paradise of God.

8 And unto the angel of the church in Smyrna write; These things saith the first and the last, which was dead, and is alive;

9 I know thy works, and tribulation, and poverty, (but thou art rich) and I know the blasphemy of them which say they are Jews, and are not, but are the synagogue of Satan.

10 Fear none of those things which thou shalt suffer: behold, the devil shall cast some of you into prison, that ye may be tried; and ye shall have tribulation ten days: be thou faithful unto death, and I will give thee a crown of life.

11 He that hath an ear, let him hear what the Spirit saith unto the churches; He that overcometh shall not be hurt of the second death.

12 And to the angel of the church in Pergamos write; These things saith he which hath the sharp sword with two edges;

13 I know thy works, and where thou dwellest, even where Satan's seat is: and thou holdest fast my name, and hast not denied my faith, even in those days wherein Antipas was my faithful martyr, who was slain among you, where Satan dwelleth.

14 But I have a few things against thee, because thou hast there them that hold the doctrine of Balaam, who taught Balac to cast a stumblingblock before the children of Israel, to eat things sacrificed unto idols, and to commit fornication.

15 So hast thou also them that hold the doctrine of the Nicolaitans, which thing I hate.

16 Repent; or else I will come unto thee quickly, and will fight against them with the sword of my mouth.

17 He that hath an ear, let him hear what the Spirit saith unto the churches; To him that overcometh will I give to eat of the hidden manna, and will give him a white stone, and in the stone a new name written, which no man knoweth saving he that receiveth it.

18 And unto the angel of the church in Thyatira write; These things saith the Son of God, who hath his eyes like unto a flame of fire, and his feet are like fine brass;

19 I know thy works, and charity, and service, and faith, and thy patience, and thy works; and the last to be more than the first.

20 Notwithstanding I have a few things against thee, because thou sufferest that woman Jezebel, which calleth herself a prophetess, to teach and to seduce my servants to commit fornication, and to eat things sacrificed unto idols.

21 And I gave her space to repent of her fornication; and she repented not.

22 Behold, I will cast her into a bed, and them that commit adultery with her into great tribulation, except they repent of their deeds.

23 And I will kill her children with death; and all the churches shall know that I am he which searcheth the reins and hearts: and I will give unto every one of you according to your works.

24 But unto you I say, and unto the rest in Thyatira, as many as have not this doctrine, and which have not known the depths of Satan, as they speak; I will put upon you none other burden.

25 But that which ye have already hold fast till I come.

26 And he that overcometh, and keepeth my works unto the end, to him will I give power over the nations:

27 And he shall rule them with a rod of iron; as the vessels of a potter shall they be broken to shivers: even as I received of my Father.

28 And I will give him the morning star.

29 He that hath an ear, let him hear what the Spirit saith unto the churches.
\appendix
\backmatter
\end{document}