%CHAPTER 11
%\chapterfigure[scale=0.2]{Durer/Durer_Revelation_Four_Riders.jpg}{The Four Horsemen. Albrecht Dürer, 1498.}

\chapter{The Two Witnesses and \linebreak[0] \mbox{the Seventh Trumpet}}
\subsection*{The Two Witnesses}
\lettrine[lines=3,slope=0.5em]{\textcolor{red}{A}}{\ nd} there was given me a reed like unto a rod:%
	\cbiblefoot{Ezekiel}{40:3, 5}{There was a man, whose appearance was like the appearance of brass, with a line of flax in his hand, and a measuring reed\ldots six cubits long, of a cubit and a handbreadth each: so he measured\ldots} %
and one said, Rise, and measure the temple of God, and the altar, and them that worship therein. 
\vnum{2}And the court which is without the temple leave without, and measure it not; for it hath been given unto the nations: and the holy city shall they tread under foot forty and two months.%
	\cbiblefootduo{Psalms}{79:1-3}{O God, the nations are come into thine inheritance; thy holy temple have they defiled; they have laid Jerusalem in heaps. The dead bodies of thy servants have they given to be food unto the birds of the heavens, the flesh of thy saints unto the beasts of the earth. Their blood have they shed like water round about Jerusalem; and there was none to bury them.}%
					{Isaiah}{63:17-19}{Return for thy servants’ sake, the tribes of thine inheritance. Thy holy people possessed it but a little while: our adversaries have trodden down thy sanctuary. We are become as they over whom thou never barest rule, as they that were not called by thy name.} %
\vnum{3}And I will give unto my two witnesses, and they shall prophesy a thousand two hundred and threescore days,%
	\cbiblefootduosb{Daniel}{7:25}{He shall speak words against the Most High, and shall wear out the saints of the Most High; and he shall think to change the times and the law; and they shall be given into his hand until a time and times and half a time}%
					{12:7}{I heard the man clothed in linen, who was above the waters of the river, when he held up his right hand and his left hand unto heaven, and sware by him that liveth for ever that it shall be for a time, times, and a half; and when they have made an end of breaking in pieces the power of the holy people, all these things shall be finished.} %
clothed in sackcloth. 
\vnum{4}These are the two olive trees and the two candlesticks, standing before the Lord of the earth.%
	\cbiblefoot{Zechariah}{4:2-3, 11-14}{Behold, a candlestick all of gold, with its bowl upon the top of it, and its seven lamps thereon; there are seven pipes to each of the lamps, which are upon the top thereof; and two olive-trees by it, one upon the right side of the bowl, and the other upon the left side thereof\ldots Then answered I, and said unto him, What are these two olive-trees upon the right side of the candlestick and upon the left side thereof? And I answered the second time, and said unto him, What are these two olive-branches, which are beside the two golden spouts, that empty the golden oil out of themselves? And he answered me and said, Knowest thou not what these are? And I said, No, my lord. Then said he, These are the two anointed ones, that stand by the Lord of the whole earth.} %
\vnum{5}And if any man desireth to hurt them, fire proceedeth out of their mouth and devoureth their enemies;%
	\footnote{\cbibleref{IISamuel}{22:9}{There went up a smoke out of [God's] nostrils, and fire out of his mouth devoured: coals were kindled by it.} %
			  \cbibleref{Jeremiah}{5:14}{Thus saith Jehovah, the God of hosts, "Because ye speak this word, behold, I will make my words in thy mouth fire, and this people wood, and it shall devour them."} %
			  \cbibleref{Psalms}{97:3}{A fire goeth before him, and burneth up his adversaries round about} %
			  \cbibleref{IIKings}{1:10}{And Elijah answered and said to the captain of fifty, If I be a man of God, let fire come down from heaven, and consume thee and thy fifty. And there came down fire from heaven, and consumed him and his fifty.}} %
and if any man shall desire to hurt them, in this manner must he be killed. 
\vnum{6}These have the power to shut the heaven, that it rain not during the days of their prophecy:%
	\cbiblefoot{IKings}{17:1}{Elijah the Tishbite, who was of the sojourners of Gilead, said unto Ahab, As Jehovah, the God of Israel, liveth, before whom I stand, there shall not be dew nor rain these years, but according to my word.} %
 and they have power over the waters to turn them into blood,%
	\cbiblefoot{Exodus}{7:17-19}{Thus saith Jehovah, In this thou shalt know that I am Jehovah: behold, I will smite with the rod that is in my hand upon the waters which are in the river, and they shall be turned to blood. And the fish that are in the river shall die, and the river shall become foul; and the Egyptians shall loathe to drink water from the river. And Jehovah said unto Moses, Say unto Aaron, Take thy rod, and stretch out thy hand over the waters of Egypt, over their rivers, over their streams, and over their pools, and over all their ponds of water, that they may become blood; and there shall be blood throughout all the land of Egypt, both in vessels of wood and in vessels of stone.} %
 and to smite the earth with every plague,%
	\cbiblefoot{ISamuel}{4:8}{Woe unto us! who shall deliver us out of the hand of these mighty gods? These are the gods that smote the Egyptians with all manner of plagues in the wilderness.} %
as often as they shall desire. 
\vnum{7}And when they shall have finished their testimony, the beast that cometh up out of the abyss%
	\cbiblefoot{Daniel}{7:3}{Four great beasts came up from the sea, diverse one from another.} %
 shall make war with them, and overcome them, and kill them.%
	\cbiblefoot{Daniel}{7:21}{I beheld, and the same horn made war with the saints, and prevailed against them.} %
\vnum{8}And their dead bodies lie in the street%
	\cbiblefoot{Ezekiel}{11:6}{Ye have multiplied your slain in this city, and ye have filled the streets thereof with the slain.} %
of the great city,%
	\cbiblefoot{Jeremiah}{22:8}{Many nations shall pass by this city, and they shall say every man to his neighbor, Wherefore hath Jehovah done thus unto this great city?} %
which spiritually is called Sodom%
	\cbiblefootduo{Isaiah}{1:9-10}{Except Jehovah of hosts had left unto us a very small remnant, we should have been as Sodom, we should have been like unto Gomorrah. Hear the word of Jehovah, ye rulers of Sodom; give ear unto the law of our God, ye people of Gomorrah}%
				   {Ezekiel}{16:46, 49}{Thine elder sister is Samaria, that dwelleth at thy left hand, she and her daughters; and thy younger sister, that dwelleth at thy right hand, is Sodom and her daughters\ldots Behold, this was the iniquity of thy sister Sodom: pride, fulness of bread, and prosperous ease was in her and in her daughters; neither did she strengthen the hand of the poor and needy.} %
and Egypt,%
	\cbiblefoot{Joel}{3:19}{Egypt shall be a desolation\ldots for the violence done to the children of Judah, because they have shed innocent blood in their land.} %
where also their Lord was crucified. 
\vnum{9}And from among the peoples and tribes and tongues and nations do men look upon their dead bodies three days and a half, and suffer not their dead bodies to be laid in a tomb. 
\vnum{10}And they that dwell on the earth rejoice over them, and make merry; and they shall send gifts one to another; because these two prophets tormented them that dwell on the earth.%
	\cbiblefoot{Psalms}{105:38}{Egypt was glad when they departed; for the fear of them had fallen upon them.} %
\vnum{11}And after the three days and a half the breath of life from God entered into them,%
	\cbiblefoot{Ezekiel}{37:5}{Thus saith the Lord Jehovah unto these bones: Behold, I will cause breath to enter into you, and ye shall live.} %
and they stood upon their feet;% 
	\cbiblefoot{Ezekiel}{37:10}{I prophesied as he commanded me, and the breath came into them, and they lived, and stood up upon their feet, an exceeding great army.} %
and great fear fell upon them%
	\cbiblefootduo{Genesis}{15:12}{When the sun was going down, a deep sleep fell upon Abram; and, lo, a horror of great darkness fell upon him}%
				  {Exodus}{15:16}{Terror and dread falleth upon them; by the greatness of thine arm they are as still as a stone; till thy people pass over, O Jehovah, till the people pass over that thou hast purchased.} %
that beheld them. 
\vnum{12}And they heard a great voice from heaven saying unto them, Come up hither. And they went up into heaven in the cloud;%
	\cbiblefoot{IIKings}{2:11}{As they still went on, and talked\ldots behold, there appeared a chariot of fire, and horses of fire, which parted them both asunder; and Elijah went up by a whirlwind into heaven.} %
and their enemies beheld them. 
\vnum{13}And in that hour there was a great earthquake,%
	\cbiblefoot{Ezekiel}{38:19-20}{In my jealousy and in the fire of my wrath have I spoken, Surely in that day there shall be a great shaking in the land of Israel; so that the fishes of the sea, and the birds of the heavens, and the beasts of the field, and all creeping things that creep upon the earth, and all the men that are upon the face of the earth, shall shake at my presence, and the mountains shall be thrown down, and the steep places shall fall, and every wall shall fall to the ground.} %
and the tenth part of the city fell; and there were killed in the earthquake seven thousand persons: and the rest were affrighted, and gave glory to the God of heaven.

\vnum{14}The second Woe is past: behold, the third Woe cometh quickly.
\subsection*{The Seventh Trumpet}
\vnum{15}And the seventh angel sounded; and there followed great voices in heaven, and they said,
\begin{verse}
The kingdom of the world is become the kingdom of our Lord,%
	\footnote{\cbibleref{Obadiah}{1:21}{Saviours shall come up on mount Zion to judge the mount of Esau; and the kingdom shall be Jehovah’s} %
			  \cbibleref{Psalms}{22:27-28}{All the ends of the earth shall remember and turn unto Jehovah; and all the kindreds of the nations shall worship before thee. For the kingdom is Jehovah’s; and he is the ruler over the nations} %
			  \cbiblechvs{Psalms}{2:2, 6, 8}{The kings of the earth set themselves, and the rulers take counsel together, against Jehovah, and against his anointed\ldots Yet I have set my king upon my holy hill of Zion\ldots Ask of me, and I will give thee the nations for thine inheritance, and the uttermost parts of the earth for thy possession.}} %
and of his Christ:%
	\cbiblefoot{Daniel}{7:14}{There was given him dominion, and glory, and a kingdom, that all the peoples, nations, and languages should serve him: his dominion is an everlasting dominion, which shall not pass away, and his kingdom that which shall not be destroyed.}\\* 
and he shall reign for ever and ever.%
	\footnote{\cbibleref{Psalms}{10:16}{Jehovah is King for ever and ever: the nations are perished out of his land} %
			  \cbibleref{Daniel}{2:44}{In the days of those kings shall the God of heaven set up a kingdom which shall never be destroyed, nor shall the sovereignty thereof be left to another people; but it shall break in pieces and consume all these kingdoms, and it shall stand for ever} %
			  \cbiblechvs{Daniel}{7:27}{The kingdom and the dominion, and the greatness of the kingdoms under the whole heaven, shall be given to the people of the saints of the Most High: his kingdom is an everlasting kingdom, and all dominions shall serve and obey him.}} %			  
\end{verse}
\vnum{16}And the four and twenty elders, who sit before God on their thrones, fell upon their faces and worshipped God, 
\vnum{17}saying,
\begin{verse}
We give thee thanks, O Lord God, the Almighty,\\*\vin
who art and who wast;\\*
because thou hast taken thy great power,\\*\vin
and didst reign.\\ 
\vleftofline{\vnum{18}}And the nations were wroth,\\*\vin
and thy wrath came,\\*\vin
and the time of the dead to be judged,\\
and the time to give their reward to thy servants the prophets,\\*\vin
and to the saints, and to them that fear thy name,\\*\vin
the small and the great;%
	\cbiblefoot{Psalms}{115:13}{He will bless them that fear Jehovah, both small and great.}\\
and to destroy them that destroy the earth.%
	\cbiblefoot{Jeremiah}{51:25}{Behold, I am against thee, O destroying mountain, saith Jehovah, which destroyest all the earth; and I will stretch out my hand upon thee, and roll thee down from the rocks, and will make thee a burnt mountain.} %
\end{verse}
\vnum{19}And there was opened the temple of God that is in heaven; and there was seen in his temple the ark of his covenant; and there followed lightnings, and voices, and thunders, and an earthquake, and great hail.%
	\cbiblefoot{Isaiah}{29:5-7}{The multitude of thy foes shall be like small dust, and the multitude of the terrible ones as chaff that passeth away: yea, it shall be in an instant suddenly. She shall be visited of Jehovah of hosts with thunder, and with earthquake, and great noise, with whirlwind and tempest, and the flame of a devouring fire. And the multitude of all the nations that fight against Ariel, even all that fight against her and her stronghold, and that distress her, shall be as a dream, a vision of the night.}
