%CHAPTER 16
%\chapterfigure[scale=0.2]{Durer/Durer_Revelation_Four_Riders.jpg}{The Four Horsemen. Albrecht Dürer, 1498.}

\chapter{The Bowls of Wrath}
\fancyhead{} %clear formatting
\fancyhead[LE]{CHAPTER 16} %for even pages, put the chapter number on left side; for odd pages, put chapter title on right
\fancyhead[RO]{THE BOWLS OF WRATH}
\lettrine[lines=3,slope=0.5em]{\textcolor{red}{A}}{\ nd I heard a great voice} out of the temple,%
	\cbiblefoot{Isaiah}{66:6}{A voice of tumult from the city, a voice from the temple, a voice of Jehovah that rendereth recompense to his enemies.} %
 saying to the seven angels, Go ye, and pour out the seven bowls of the wrath of God into the earth.%
	\footnote{\cbibleref{Jeremiah}{10:25}{Pour out thy wrath upon the nations that know thee not, and upon the families that call not on thy name; for they have devoured Jacob, yea, they have devoured him and consumed him, and have laid waste his habitation} %
			  \cbibleref{Ezekiel}{14:19}{Or if I send a pestilence into that land, and pour out my wrath upon it in blood, to cut off from it man and beast} %
			  \cbibleref{Zephaniah}{3:8}{Wait ye for me, saith Jehovah, until the day that I rise up to the prey; for my determination is to gather the nations, that I may assemble the kingdoms, to pour upon them mine indignation, even all my fierce anger; for all the earth shall be devoured with the fire of my jealousy} %
			  \cbibleref{Psalms}{69:24}{Pour out thine indignation upon them, and let the fierceness of thine anger overtake them.}}
\newline

\zz
\vnum{2}And the first went, and poured out his bowl into the earth; and it became a noisome and grievous sore%
	\cbiblefootduo{Exodus}{9:10-11}{They took ashes of the furnace, and stood before Pharaoh; and Moses sprinkled it up toward heaven; and it became a boil breaking forth with blains upon man and upon beast. And the magicians could not stand before Moses because of the boils; for the boils were upon the magicians, and upon all the Egyptians}%
				   {Deuteronomy}{28:35}{Jehovah will smite thee in the knees, and in the legs, with a sore boil, whereof thou canst not be healed, from the sole of thy foot unto the crown of thy head.} %
upon the men that had the mark of the beast, and that worshipped his image.

\vnum{3}And the second poured out his bowl into the sea; and it became blood%
	\cbiblefootduo{Exodus}{7:17-21}{Thus saith Jehovah, In this thou shalt know that I am Jehovah: behold, I will smite with the rod that is in my hand upon the waters which are in the river, and they shall be turned to blood. And the fish that are in the river shall die, and the river shall become foul; and the Egyptians shall loathe to drink water from the river. And Jehovah said unto Moses, Say unto Aaron, Take thy rod, and stretch out thy hand over the waters of Egypt, over their rivers, over their streams, and over their pools, and over all their ponds of water, that they may become blood; and there shall be blood throughout all the land of Egypt, both in vessels of wood and in vessels of stone. And Moses and Aaron did so, as Jehovah commanded; and he lifted up the rod, and smote the waters that were in the river, in the sight of Pharaoh, and in the sight of his servants; and all the waters that were in the river were turned to blood. And the fish that were in the river died; and the river became foul, and the Egyptians could not drink water from the river; and the blood was throughout all the land of Egypt}%
				  {Psalms}{78:44}{And turned their rivers into blood, and their streams, so that they could not drink\ldots} %
as of a dead man; and every living soul died, even the things that were in the sea.

\vnum{4}And the third poured out his bowl into the rivers and the fountains of the waters; and it became blood. 
\vnum{5}And I heard the angel of the waters saying, 
\begin{verse}
Righteous art thou,%
	\cbiblefootduo{Deuteronomy}{32:4}{The Rock, his work is perfect; for all his ways are justice: a God of faithfulness and without iniquity, just and right is he}%
				  {Psalms}{145:17}{Jehovah is righteous in all his ways, and gracious in all his works.} %
who art and who wast, thou Holy One,\\*\vin because thou didst thus judge:%
	\cbiblefoot{Psalms}{119:137}{Jehovah is righteous in all his ways, and gracious in all his works.}\\
\vleftofline{\vnum{6}}for they poured out the blood of saints and prophets,%
	\cbiblefoot{Psalms}{79:3}{Their blood have they shed like water round about Jerusalem; and there was none to bury them.}\\*\vin
 and blood hast thou given them to drink:%
	\cbiblefoot{Isaiah}{49:26}{I will feed them that oppress thee with their own flesh; and they shall be drunken with their own blood, as with sweet wine: and all flesh shall know that I, Jehovah, am thy Saviour, and thy Redeemer, the Mighty One of Jacob.}\\*
 they are worthy. 
 \end{verse}
\vnum{7}And I heard the altar saying, Yea, O Lord God, the Almighty, true and righteous are thy judgments.%
	\cbiblefootduosb{Psalms}{19:9}{The fear of Jehovah is clean, enduring for ever: the ordinances of Jehovah are true, and righteous altogether}%
				     {119:137}{Righteous art thou, O Jehovah, and upright are thy judgments.} %

\vnum{8}And the fourth poured out his bowl upon the sun; and it was given unto it to scorch men with fire. 
\vnum{9}And men were scorched with great heat: and they blasphemed%
	\cbiblefootduosb{Isaiah}{8:21}{They shall pass through it, sore distressed and hungry; and it shall come to pass that, when they shall be hungry, they shall fret themselves, and curse by their king and by their God, and turn their faces upward}%
					{52:5}{What do I here, saith Jehovah, seeing that my people is taken away for nought? They that rule over them do howl, saith Jehovah, and my name continually all the day is blasphemed.} %
the name of God who hath the power over these plagues; and they repented not to give him glory.

\vnum{10}And the fifth poured out his bowl upon the throne of the beast; and his kingdom was darkened;%
	\cbiblefootduo{Exodus}{10:21-22}{Jehovah said unto Moses, Stretch out thy hand toward heaven, that there may be darkness over the land of Egypt, even darkness which may be felt. And Moses stretched forth his hand toward heaven; and there was a thick darkness in all the land of Egypt three days; they saw not one another, neither rose any one from his place for three days: but all the children of Israel had light in their dwellings}%
				  {Isaiah}{8:22}{They shall look unto the earth, and, behold, distress and darkness, the gloom of anguish; and into thick darkness they shall be driven away.} %
and they gnawed their tongues for pain, 
\vnum{11}and they blasphemed the God of heaven%
	\cbiblefoot{Daniel}{2:18-19}{That they would desire mercies of the God of heaven concerning this secret; that Daniel and his companions should not perish with the rest of the wise men of Babylon. Then was the secret revealed unto Daniel in a vision of the night. Then Daniel blessed the God of heaven.} %
because of their pains and their sores; and they repented not of their works.

\vnum{12}And the sixth poured out his bowl upon the great river, the river Euphrates; and the water thereof was dried up,%
	\footnote{\cbibleref{Isaiah}{11:15}{Jehovah will utterly destroy the tongue of the Egyptian sea; and with his scorching wind will he wave his hand over the [Eufrates] River, and will smite it into seven streams, and cause men to march over dryshod} %
				\cbiblechvs{Isaiah}{44:27}{[I am Jehovah] that saith to the deep, Be dry, and I will dry up thy rivers} %
				\cbibleref{Jeremiah}{50:38}{A drought is upon her [Babylon's] waters, and they shall be dried up; for it is a land of graven images, and they are mad over idols} %
				\cbiblechvs{Jeremiah}{51:36}{Thus saith Jehovah: Behold, I will plead thy cause, and take vengeance for thee; and I will dry up her [Babylon's] sea, and make her fountain dry.}} %
that the way might be made ready%
	\footnote{\cbibleref{Isaiah}{11:16}{There shall be a highway for the remnant of his people, that shall remain, from Assyria; like as there was for Israel in the day that he came up out of the land of Egypt} %
			  \cbiblechvs{Isaiah}{51:10}{Is it not thou that driedst up the sea, the waters of the great deep; that madest the depths of the sea a way for the redeemed to pass over?} %
			  \cbibleref{Psalms}{106:9}{He rebuked the Red Sea also, and it was dried up: So he led them through the depths, as through a wilderness.}} %
for the kings that come from the sunrising.%
	\footnote{\cbibleref{Isaiah}{41:2}{Who hath raised up one from the east, whom he calleth in righteousness to his foot? he giveth nations before him, and maketh him rule over kings; he giveth them as the dust to his sword, as the driven stubble to his bow} %
			  \cbiblevs{Isaiah}{41:25}{I have raised up one from the north, and he is come; from the rising of the sun one that calleth upon my name: and he shall come upon rulers as upon mortar, and as the potter treadeth clay} %
			  \cbiblechvs{Isaiah}{46:9-11}{Remember the former things of old: for I am God, and there is none else; I am God, and there is none like me; declaring the end from the beginning, and from ancient times things that are not yet done; saying, My counsel shall stand, and I will do all my pleasure; calling a ravenous bird from the east, the man of my counsel from a far country; yea, I have spoken, I will also bring it to pass; I have purposed, I will also do it.}} %
\vnum{13}And I saw coming out of the mouth of the dragon, and out of the mouth of the beast, and out of the mouth of the false prophet, three unclean spirits,%
	\cbiblefoot{IKings}{22:21-23}{There came forth a spirit, and stood before Jehovah, and said, I will entice him. And Jehovah said unto him, Wherewith? And he said, I will go forth, and will be a lying spirit in the mouth of all his prophets. And he said, Thou shalt entice him, and shalt prevail also: go forth, and do so. Now therefore, behold, Jehovah hath put a lying spirit in the mouth of all these thy prophets; and Jehovah hath spoken evil concerning thee.} %
as it were frogs:%
	\footnote{\cbibleref{Exodus}{8:3}{The river shall swarm with frogs, which shall go up and come into thy house, and into thy bedchamber, and upon thy bed, and into the house of thy servants, and upon thy people, and into thine ovens, and into thy kneading-troughs} %
					\cbibleref{Psalms}{78:45}{He sent among them swarms of flies, which devoured them; and frogs, which destroyed them} %
					\cbiblechvs{Psalms}{105:30}{Their land swarmed with frogs in the chambers of their kings.}} %
\vnum{14}for they are spirits of demons, working signs; which go forth unto the kings of the whole world, to gather them together unto the war of the great day of God, the Almighty. 
\vnum{15}(Behold, I come as a thief. Blessed is he that watcheth, and keepeth his garments, lest he walk naked, and they see his shame.) 
\vnum{16}And they gathered them together into the place which is called in Hebrew Har-Magedon.%
	\footnote{\cbibleref{IIKings}{9:27}{When Ahaziah the king of Judah saw this, he fled by the way of the garden-house. And Jehu followed after him, and said, Smite him also in the chariot: and they smote him at the ascent of Gur, which is by Ibleam. And he fled to Megiddo, and died there} %
				\cbiblechvs{IIKings}{23:29}{In his days Pharaoh-necoh king of Egypt went up against the king of Assyria to the river Euphrates: and king Josiah went against him; and Pharaoh-necoh slew him at Megiddo, when he had seen him} %
				\cbibleref{Zechariah}{12:11}{In that day shall there be a great mourning in Jerusalem, as the mourning of Hadadrimmon in the valley of Megiddon.}} %

\vnum{17}And the seventh poured out his bowl upon the air; and there came forth a great voice out of the temple,%
	\cbiblefoot{Isaiah}{66:6}{A voice of tumult from the city, a voice from the temple, a voice of Jehovah that rendereth recompense to his enemies.} %
from the throne, saying, It is done: 
\vnum{18}and there were lightnings, and voices, and thunders; and there was a great earthquake, such as was not since there were men upon the earth,%
	\cbiblefootduo{Daniel}{12:1}{At that time shall Michael stand up, the great prince who standeth for the children of thy people; and there shall be a time of trouble, such as never was since there was a nation even to that same time: and at that time thy people shall be delivered, every one that shall be found written in the book}%
					{Exodus}{9:24}{There was hail, and fire mingled with the hail, very grievous, such as had not been in all the land of Egypt since it became a nation.} %
so great an earthquake, so mighty. 
\vnum{19}And the great city was divided into three parts, and the cities of the nations fell: and Babylon the great was remembered in the sight of God, to give unto her the cup of the wine of the fierceness of his wrath. 
\vnum{20}And every island fled away, and the mountains were not found. 
\vnum{21}And great hail, every stone about the weight of a talent, cometh down out of heaven upon men: and men blasphemed God because of the plague of the hail; for the plague thereof is exceeding great.
