%CHAPTER 3
\chapter{Letters to Churches II}
\subsection*{Letter to Sardis}
\lettrine[lines=3]{\textcolor{red}{A}}{\ nd to the angel of the church in Sardis write:}

\zz These things saith he that hath the seven Spirits of God, and the seven stars: I know thy works, that thou hast a name that thou livest, and thou art dead. %
\vnum{2}Be thou watchful, and establish the things that remain, which were ready to die: for I have found no works of thine perfected before my God. %
\vnum{3}Remember therefore how thou hast received and didst hear; and keep it, and repent. If therefore thou shalt not watch, I will come as a thief, and thou shalt not know what hour I will come upon thee. %
\vnum{4}But thou hast a few names in Sardis that did not defile their garments: and they shall walk with me in white; for they are worthy. %
\vnum{5}He that overcometh shall thus be arrayed in white garments; and I will in no wise blot his name out of the book of life,%
	\footnote{\cbibleref{Exodus}{32:32-33}{Yet now, if thou wilt forgive their sin—; and if not, blot me, I pray thee, out of thy book which thou hast written. And Jehovah said unto Moses, Whosoever hath sinned against me, him will I blot out of my book.} %
			\cbibleref{Psalms}{69:28}{Let them be blotted out of the book of life, And not be written with the righteous.} %
			\cbibleref{Daniel}{12:1}{Thy people shall be delivered, every one that shall be found written in the book.} %
			\cbibleref{Malachi}{3:16}{Then they that feared Jehovah spake one with another; and Jehovah hearkened, and heard, and a book of remembrance was written before him, for them that feared Jehovah, and that thought upon his name.}} %
and I will confess his name before my Father, and before his angels. %
\vnum{6}He that hath an ear, let him hear what the Spirit saith to the churches.
\subsection*{Letter to Philadelphia}
\vnum{7}And to the angel of the church in Philadelphia write:

These things saith he that is holy, he that is true, he that hath the key of David, he that openeth and none shall shut, and that shutteth and none openeth:%
	\cbiblefootduo{Isaiah}{22:20-23}{It shall come to pass in that day, that I will call my servant Eliakim\ldots and I will commit thy government into his hand; and he shall be a father to the inhabitants of Jerusalem, and to the house of Judah. And the key of the house of David will I lay upon his shoulder; and he shall open, and none shall shut; and he shall shut, and none shall open. And I will fasten him as a nail in a sure place; and he shall be for a throne of glory to his father’s house.}%
				{Job}{12:14}{With God is wisdom and might\ldots Behold, he breaketh down, and it cannot be built again; He shutteth up a man, and there can be no opening.} %
\vnum{8}I know thy works (behold, I have set before thee a door opened, which none can shut), that thou hast a little power, and didst keep my word, and didst not deny my name. %
\vnum{9} Behold, I give of the synagogue of Satan, of them that say they are Jews, and they are not, but do lie; behold, I will make them to come and worship before thy feet,%
	\footnote{\cbibleref{Isaiah}{49:23}{Kings shall be thy nursing fathers, and their queens thy nursing mothers: they shall bow down to thee with their faces to the earth, and lick the dust of thy feet; and thou shalt know that I am Jehovah; and they that wait for me shall not be put to shame} %
				\cbiblechvs{Isaiah}{60:14}{The sons of them that afflicted thee shall come bending unto thee; and all they that despised thee shall bow themselves down at the soles of thy feet; and they shall call thee The city of Jehovah, The Zion of the Holy One of Israel.}} %
and to know that I have loved thee.%
 	\cbiblefoot{Isaiah}{43:4}{Since thou hast been precious in my sight, and honorable, and I have loved thee; therefore will I give men in thy stead, and peoples instead of thy life.} %
\vnum{10}Because thou didst keep the word of my patience, I also will keep thee from the hour of trial, that hour which is to come upon the whole world, to try them that dwell upon the earth. %
\vnum{11}I come quickly: hold fast that which thou hast, that no one take thy crown. %
\vnum{12}He that overcometh, I will make him a pillar in the temple of my God, and he shall go out thence no more: and I will write upon him the name of my God, and the name of the city of my God,%
	\cbiblefoot{Ezekiel}{48:35}{The name of the city from that day shall be, Jehovah is there.}
 the new Jerusalem, which cometh down out of heaven from my God, and mine own new name. %
\vnum{13}He that hath an ear, let him hear what the Spirit saith to the churches.
\subsection*{Letter to Laodicea}
\vnum{14}And to the angel of the church in Laodicea write:

These things saith the Amen, the faithful and true witness,%
	\cbiblefoot{Isaiah}{65:16}{He who blesseth himself in the earth shall bless himself in the God of truth; and he that sweareth in the earth shall swear by the God of truth.} %
 the beginning of the creation of God:%
 	\cbiblefoot{Proverbs}{8:22-23}{Jehovah possessed me in the beginning of his way, before his works of old. I was set up from everlasting, from the beginning, before the earth was.} %
\vnum{15}I know thy works, that thou art neither cold nor hot: I would thou wert cold or hot. %
\vnum{16}So because thou art lukewarm, and neither hot nor cold, I will spew thee out of my mouth. %
\vnum{17}Because thou sayest, I am rich,%
	\cbiblefootduo{Zechariah}{11:4-5}{Thus said Jehovah my God: Feed the flock of slaughter; whose possessors slay them, and hold themselves not guilty; and they that sell them say, Blessed be Jehovah, for I am rich; and their own shepherds pity them not.}	
				{Hosea}{12:8-9}{Ephraim said, Surely I am become rich, I have found me wealth: in all my labors they shall find in me no iniquity that were sin. But I am Jehovah thy God from the land of Egypt; I will yet again make thee to dwell in tents, as in the days of the solemn feast.} %
and have gotten riches, and have need of nothing; and knowest not that thou art the wretched one and miserable and poor and blind and naked: %
\vnum{18}I counsel thee to buy of me gold refined by fire, that thou mayest become rich; and white garments, that thou mayest clothe thyself, and that the shame of thy nakedness be not made manifest; and eyesalve to anoint thine eyes, that thou mayest see.%
	\cbiblefoot{Isaiah}{55:1-3}{Ho, every one that thirsteth, come ye to the waters, and he that hath no money; come ye, buy, and eat; yea, come, buy wine and milk without money and without price. Wherefore do ye spend money for that which is not bread? and your labor for that which satisfieth not? hearken diligently unto me, and eat ye that which is good, and let your soul delight itself in fatness. Incline your ear, and come unto me; hear, and your soul shall live: and I will make an everlasting covenant with you, even the sure mercies of David.} %
\vnum{19}As many as I love, I reprove and chasten:%
	\cbiblefoot{Proverbs}{3:11-12}{My son, despise not the chastening of Jehovah; neither be weary of his reproof: for whom Jehovah loveth he reproveth, even as a father the son in whom he delighteth.} %
be zealous therefore, and repent. %
\vnum{20}Behold, I stand at the door and knock:%
	\cbiblefoot{SongofSongs}{5:2}{I was asleep, but my heart waked: it is the voice of my beloved that knocketh, saying, open to me, my sister, my love, my dove, my undefiled; for my head is filled with dew, my locks with the drops of the night.} %
if any man hear my voice and open the door, I will come in to him, and will sup with him, and he with me. %
\vnum{21}He that overcometh, I will give to him to sit down with me in my throne, as I also overcame, and sat down with my Father in his throne. %
\vnum{22}He that hath an ear, let him hear what the Spirit saith to the churches.
