%CHAPTER 21
%\chapterfigure[scale=0.2]{Durer/Durer_Revelation_Four_Riders.jpg}{The Four Horsemen. Albrecht Dürer, 1498.}

\chapter{The New Heaven and Earth}
\lettrine[lines=3,slope=0.5em]{\textcolor{red}{A}}{\ nd} I saw a new heaven and a new earth:%
	\cbiblefootduosb{Isaiah}{65:17}{For, behold, I create new heavens and a new earth; and the former things shall not be remembered, nor come into mind}%
				{66:22}{For as the new heavens and the new earth, which I will make, shall remain before me, saith Jehovah, so shall your seed and your name remain.} %
for the first heaven and the first earth are passed away; and the sea is no more. 
\vnum{2}And I saw the holy city,%
	\cbiblefootduosb{Nehemiah}{11:1}{The princes of the people dwelt in Jerusalem: the rest of the people also cast lots, to bring one of ten to dwell in Jerusalem the holy city, and nine parts in the other cities}%
					{11:18}{All the Levites in the holy city\ldots} %
new Jerusalem, coming down out of heaven from God, made ready as a bride adorned for her husband.%
	\cbiblefoot{Isaiah}{61:10}{I will greatly rejoice in Jehovah, my soul shall be joyful in my God; for he hath clothed me with the garments of salvation, he hath covered me with the robe of righteousness, as a bridegroom decketh himself with a garland, and as a bride adorneth herself with her jewels.} %
\vnum{3}And I heard a great voice out of the throne saying,
\begin{verse}
Behold, the tabernacle of God is with men,\\*\vin
and he shall dwell with them,%
	\footnote{
		\cbibleref{Zechariah}{2:10}{Sing and rejoice, O daughter of Zion; for, lo, I come, and I will dwell in the midst of thee, saith Jehovah} %
		\cbibleref{Leviticus}{26:12}{And I will walk among you, and will be your God, and ye shall be my people} %
		\cbibleref{Ezekiel}{37:27}{My tabernacle also shall be with them; and I will be their God, and they shall be my people.}%
	}\\*\vin
and they shall be his peoples,%
	\footnote{
		\cbibleref{Jeremiah}{31:1}{At that time, saith Jehovah, will I be the God of all the families of Israel, and they shall be my people} %
		\cbibleref{Ezekiel}{11:20}{That they may walk in my statutes, and keep mine ordinances, and do them: and they shall be my people, and I will be their God} %
		\cbibleref{Zechariah}{8:8}{I will bring them, and they shall dwell in the midst of Jerusalem; and they shall be my people, and I will be their God, in truth and in righteousness} %
		\cbibleref{Psalms}{95:7}{He is our God, and we are the people of his pasture, and the sheep of his hand.} %		
	}\\*\vin
and God himself shall be with them, and be their God:\\
\vleftofline{\vnum{4}}and he shall wipe away every tear from their eyes;%
	\cbiblefootduo{Isaiah}{25:8}{He hath swallowed up death for ever; and the Lord Jehovah will wipe away tears from off all faces; and the reproach of his people will he take away from off all the earth: for Jehovah hath spoken it}%
	{Jeremiah}{31:16}{Thus saith Jehovah: Refrain thy voice from weeping, and thine eyes from tears; for thy work shall be rewarded, saith Jehovah; and they shall come again from the land of the enemy.}\\*\vin
and death shall be no more;\\*\vin
neither shall there be mourning, nor crying, nor pain, any more:%
	\footnote{\cbibleref{Isaiah}{35:10}{The ransomed of Jehovah shall return, and come with singing unto Zion; and everlasting joy shall be upon their heads: they shall obtain gladness and joy, and sorrow and sighing shall flee away} %
		      \cbiblechvs{Isaiah}{65:19}{I will rejoice in Jerusalem, and joy in my people; and there shall be heard in her no more the voice of weeping and the voice of crying.}}\\*\vin
the first things are passed away.%
	\footnote{\cbibleref{Isaiah}{43:18}{Remember ye not the former things, neither consider the things of old} %
		      cf. \cbiblerefonly{Isaiah}{65:17} cited above.}
\end{verse}

\vnum{5}And he that sitteth on the throne said, Behold, I make all things new.%
	\cbiblefoot{Isaiah}{43:19}{Behold, I will do a new thing; now shall it spring forth; shall ye not know it? I will even make a way in the wilderness, and rivers in the desert.} %
And he saith, Write: for these words are faithful and true. 
\vnum{6}And he said unto me, They are come to pass. I am the Alpha and the Omega, the beginning and the end. I will give unto him that is athirst of the fountain of the water of life freely.%
	\footnote{\cbibleref{Isaiah}{55:1}{Ho, every one that thirsteth, come ye to the waters, and he that hath no money; come ye, buy, and eat; yea, come, buy wine and milk without money and without price} %
			  \cbibleref{Jeremiah}{2:13}{For my people have committed two evils: they have forsaken me, the fountain of living waters, and hewed them out cisterns, broken cisterns, that can hold no water} %
			  \cbibleref{Zechariah}{14:8}{It shall come to pass in that day, that living waters shall go out from Jerusalem; half of them toward the eastern sea, and half of them toward the western sea: in summer and in winter shall it be.}} %
\vnum{7}He that overcometh shall inherit these things; and I will be his God, and he shall be my son.%
	\cbiblefoot{IISamuel}{7:14}{I will be his father, and he shall be my son: if he commit iniquity, I will chasten him with the rod of men, and with the stripes of the children of men.} %
\vnum{8}But for the fearful, and unbelieving, and abominable, and murderers, and fornicators, and sorcerers, and idolaters, and all liars, their part shall be in the lake that burneth with fire and brimstone; which is the second death.

\vnum{9}And there came one of the seven angels who had the seven bowls, who were laden with the seven last plagues; and he spake with me, saying, Come hither, I will show thee the bride, the wife of the Lamb. 
\vnum{10}And he carried me away in the Spirit to a mountain great and high,%
	\cbiblefoot{Ezekiel}{40:1-2}{The hand of Jehovah was upon me, and he brought me thither. In the visions of God brought he me into the land of Israel, and set me down upon a very high mountain, whereon was as it were the frame of a city on the south.} %
and showed me the holy city Jerusalem, coming down out of heaven from God, 
\vnum{11}having the glory of God:%
	\footnote{\cbibleref{Isaiah}{58:8}{Then shall thy light break forth as the morning, and thy healing shall spring forth speedily; and thy righteousness shall go before thee; the glory of Jehovah shall be thy rearward} %
				\cbiblechvs{Isaiah}{60:1-2}{Arise, shine; for thy light is come, and the glory of Jehovah is risen upon thee. For, behold, darkness shall cover the earth, and gross darkness the peoples; but Jehovah will arise upon thee, and his glory shall be seen upon thee} %
				\cbibleref{Ezekiel}{43:2-5}{And, behold, the glory of the God of Israel came from the way of the east: and his voice was like the sound of many waters; and the earth shined with his glory. And it was according to the appearance of the vision which I saw, even according to the vision that I saw when I came to destroy the city; and the visions were like the vision that I saw by the river Chebar; and I fell upon my face. And the glory of Jehovah came into the house by the way of the gate whose prospect is toward the east. And the Spirit took me up, and brought me into the inner court; and, behold, the glory of Jehovah filled the house.}} %
her light was like unto a stone most precious, as it were a jasper stone, clear as crystal: 
\vnum{12}having a wall great and high; having twelve gates,%
	\cbiblefoot{Ezekiel}{48:30-34}{These are the egresses of the city\ldots and the gates of the city shall be after the names of the tribes of Israel, three gates northward: the gate of Reuben, one; the gate of Judah, one; the gate of Levi, one. And at the east side\ldots three gates: even the gate of Joseph, one; the gate of Benjamin, one; the gate of Dan, one. And at the south side\ldots three gates: the gate of Simeon, one; the gate of Issachar, one; the gate of Zebulun, one. At the west side\ldots three gates: the gate of Gad, one; the gate of Asher, one; the gate of Naphtali, one.} %
 and at the gates twelve angels;%
	\cbiblefoot{Isaiah}{62:6}{I have set watchmen upon thy walls, O Jerusalem; they shall never hold their peace day nor night: ye that are Jehovah’s remembrancers, take ye no rest.} %
 and names written thereon, which are the names of the twelve tribes of the children of Israel:%
	\cbiblefoot{Exodus}{28:21}{The stones shall be according to the names of the children of Israel, twelve, according to their names; like the engravings of a signet, every one according to his name, they shall be for the twelve tribes.} %
\vnum{13}on the east were three gates; and on the north three gates; and on the south three gates; and on the west three gates.
\vnum{14}And the wall of the city had twelve foundations, and on them twelve names of the twelve apostles of the Lamb. 
\vnum{15}And he that spake with me had for a measure a golden reed to measure the city, and the gates thereof, and the wall thereof.%
	\cbiblefoot{Ezekiel}{40:3, 5}{There was a man, whose appearance was like the appearance of brass, with a line of flax in his hand, and a measuring reed\ldots six cubits long, of a cubit and a handbreadth each: so he measured\ldots} %
\vnum{16}And the city lieth foursquare, and the length thereof is as great as the breadth:%
	\cbiblefoot{Ezekiel}{43:16}{The altar hearth shall be twelve cubits long by twelve broad, square in the four sides thereof.} %
and he measured the city with the reed, twelve thousand furlongs: the length and the breadth and the height thereof are equal.%
	\cbiblefoot{Ezekiel}{48:16}{These shall be the measures thereof: the north side four thousand and five hundred, and the south side four thousand and five hundred, and on the east side four thousand and five hundred, and the west side four thousand and five hundred} %
\vnum{17}And he measured the wall thereof, a hundred and forty and four cubits, according to the measure of a man, that is, of an angel. 
\vnum{18}And the building of the wall thereof was jasper: and the city was pure gold, like unto pure glass. 
\vnum{19}The foundations of the wall of the city were adorned with all manner of precious stones.%
	\cbiblefoottrio{Exodus}{28:17-20}{Thou shalt set in it settings of stones, four rows of stones: a row of sardius, topaz, and carbuncle shall be the first row; and the second row an emerald, a sapphire, and a diamond; and the third row a jacinth, an agate, and an amethyst; nd the fourth row a beryl, and an onyx, and a jasper: they shall be inclosed in gold in their settings}%
				{Isaiah}{54:11-12}{O thou afflicted, tossed with tempest, and not comforted, behold, I will set thy stones in fair colors, and lay thy foundations with sapphires. And I will make thy pinnacles of rubies, and thy gates of carbuncles, and all thy border of precious stones}%
				{Ezekiel}{28:12-14}{Thou\ldots perfect in beauty. Thou wast in Eden, the garden of God; every precious stone was thy covering, the sardius, the topaz, and the diamond, the beryl, the onyx, and the jasper, the sapphire, the emerald, and the carbuncle, and gold: the workmanship of thy tabrets and of thy pipes was in thee; in the day that thou wast created they were prepared\ldots I set thee, so that thou wast upon the holy mountain of God; thou hast walked up and down in the midst of the stones of fire.} %
 The first foundation was jasper; the second, sapphire; the third, chalcedony; the fourth, emerald; 
\vnum{20}the fifth, sardonyx; the sixth, sardius; the seventh, chrysolite; the eighth, beryl; the ninth, topaz; the tenth, chrysoprase; the eleventh, jacinth; the twelfth, amethyst. 
\vnum{21} And the twelve gates were twelve pearls; each one of the several gates was of one pearl: and the street of the city was pure gold, as it were transparent glass. 
\vnum{22}And I saw no temple therein: for the Lord God the Almighty, and the Lamb, are the temple thereof. 
\vnum{23}And the city hath no need of the sun, neither of the moon, to shine upon it: for the glory of God did lighten it, and the lamp thereof is the Lamb.%
	\footnote{\cbibleref{Isaiah}{60:1}{Arise, shine; for thy light is come, and the glory of Jehovah is risen upon thee} %
	\cbiblevs{Isaiah}{60:19-20}{The sun shall be no more thy light by day; neither for brightness shall the moon give light unto thee: but Jehovah will be unto thee an everlasting light, and thy God thy glory. Thy sun shall no more go down, neither shall thy moon withdraw itself; for Jehovah will be thine everlasting light, and the days of thy mourning shall be ended} %
	\cbiblechvs{Isaiah}{24:23}{Then the moon shall be confounded, and the sun ashamed; for Jehovah of hosts will reign in mount Zion, and in Jerusalem; and before his elders shall be glory.}} %
\vnum{24}And the nations shall walk amidst the light thereof:%
	\cbiblefoot{Isaiah}{60:3}{Nations shall come to thy light, and kings to the brightness of thy rising.} %
 and the kings of the earth bring their glory into it.
	\footnote{\cbibleref{Isaiah}{60:5}{Then thou shalt see and be radiant, and thy heart shall thrill and be enlarged; because the abundance of the sea shall be turned unto thee, the wealth of the nations shall come unto thee} %
			\cbibleref{Psalms}{68:29-32}{Because of thy temple at Jerusalem kings shall bring presents unto thee\ldots Princes shall come out of Egypt; Ethiopia shall haste to stretch out her hands unto God. Sing unto God, ye kingdoms of the earth; Oh sing praises unto the Lord} %
			\cbibleref{Psalms}{72:10-11}{The kings of Tarshish and of the isles shall render tribute: the kings of Sheba and Seba shall offer gifts. Yea, all kings shall fall down before him; all nations shall serve him.}} %
\vnum{25}And the gates thereof shall in no wise be shut by day%
	\cbiblefoot{Isaiah}{60:11}{Thy gates also shall be open continually; they shall not be shut day nor night; that men may bring unto thee the wealth of the nations, and their kings led captive.} %
(for there shall be no night there):%	
	\cbiblefoot{Zechariah}{14:7}{It shall be one day which is known unto Jehovah; not day, and not night; but it shall come to pass, that at evening time there shall be light.} %
\vnum{26}and they shall bring the glory and the honor of the nations into it: 
\vnum{27}and there shall in no wise enter into it anything unclean, or he that maketh an abomination and a lie:%
	\cbiblefootduosb{Isaiah}{35:8}{A highway shall be there, and a way, and it shall be called The way of holiness; the unclean shall not pass over it; but it shall be for the redeemed: the wayfaring men, yea fools, shall not err therein}%
				{52:1}{Awake, awake, put on thy strength, O Zion; put on thy beautiful garments, O Jerusalem, the holy city: for henceforth there shall no more come into thee the uncircumcised and the unclean.} %
but only they that are written in the Lamb’s book of life.%
	\cbiblefoot{Isaiah}{4:3}{He that is left in Zion, and he that remaineth in Jerusalem, shall be called holy, even every one that is written among the living in Jerusalem.}
