%CHAPTER 2
\chapter{Letters to Churches I}
\subsection*{Letter to Ephesus}
\lettrine[lines=3]{\textcolor{red}{T}}{\ o} the angel of the church in Ephesus write:
\zz These things saith he that holdeth the seven stars in his right hand, he that walketh in the midst of the seven golden candlesticks:%
	\cbiblefoot{Ezekiel}{28:14}{Thou wast the anointed cherub that covereth: and I set thee, so that thou wast upon the holy mountain of God; thou hast walked up and down in the midst of the stones of fire.} %
\vnum{2}I know thy works, and thy toil and patience, and that thou canst not bear evil men, and didst try them that call themselves apostles, and they are not, and didst find them false; %
\vnum{3}and thou hast patience and didst bear for my name’s sake, and hast not grown weary. %
\vnum{4}But I have this against thee, that thou didst leave thy first love.%
	\cbiblefoot{Jeremiah}{2:2-5}{Go, and cry in the ears of Jerusalem, saying, Thus saith Jehovah, I remember for thee the kindness of thy youth, the love of thine espousals; how thou wentest after me in the wilderness, in a land that was not sown. Israel was holiness unto Jehovah, the first-fruits of his increase: all that devour him shall be held guilty; evil shall come upon them, saith Jehovah\ldots Thus saith Jehovah, What unrighteousness have your fathers found in me, that they are gone far from me, and have walked after vanity, and are become vain?} %
\vnum{5}Remember%
	\cbiblefoottriosb{Ezekiel}{16:61-63}{Then shalt thou remember thy ways, and be ashamed\ldots And I will establish my covenant with thee; and thou shalt know that I am Jehovah; that thou mayest remember, and be confounded, and never open thy mouth any more, because of thy shame, when I have forgiven thee all that thou hast done, saith the Lord Jehovah}%
				{20:43}{There shall ye remember your ways, and all your doings, wherein ye have polluted yourselves; and ye shall loathe yourselves in your own sight for all your evils that ye have committed}%
				{36:31}{Then shall ye remember your evil ways, and your doings that were not good; and ye shall loathe yourselves in your own sight for your iniquities and for your abominations.} %
therefore whence thou art fallen, and repent and do the first works; or else I come to thee, and will move thy candlestick out of its place, except thou repent. %
\vnum{6}But this thou hast, that thou hatest the works of the Nicolaitans, which I also hate.%
	\cbiblefoottriosb{Psalms}{26:5}{I hate the assembly of evil-doers, and will not sit with the wicked}%
			{101:3}{I hate the work of them that turn aside; it shall not cleave unto me}%
			{139:21}{Do not I hate them, O Jehovah, that hate thee? And am not I grieved with those that rise up against thee?} %
\vnum{7}He that hath an ear, let him hear what the Spirit saith to the churches. To him that overcometh, to him will I give to eat of the tree of life, which is in the Paradise of God.%
	\cbiblefootduosb{Genesis}{2:9}{Out of the ground made Jehovah God to grow every tree that is pleasant to the sight, and good for food; the tree of life also in the midst of the garden}%
				{3:22-24}{Jehovah God said, Behold, the man is become as one of us, to know good and evil; and now, lest he put forth his hand, and take also of the tree of life, and eat, and live for ever— therefore Jehovah God sent him forth from the garden of Eden, to till the ground from whence he was taken. So he drove out the man; and he placed at the east of the garden of Eden the Cherubim, and the flame of a sword which turned every way, to keep the way of the tree of life.}
\subsection*{Letter to Smyrna}
\vnum{8}And to the angel of the church in Smyrna write:

These things saith the first and the last, who was dead, and lived again: %
\vnum{9}I know thy tribulation, and thy poverty (but thou art rich), and the blasphemy of them that say they are Jews, and they are not, but are a synagogue of Satan. %
\vnum{10}Fear not the things which thou art about to suffer: behold, the devil is about to cast some of you into prison, that ye may be tried; and ye shall have tribulation ten days.%
	\cbiblefoot{Daniel}{1:12, 14}{Prove thy servants, I beseech thee, ten days; and let them give us pulse to eat, and water to drink\ldots So he hearkened unto them in this matter, and proved them ten days.} %
Be thou faithful unto death, and I will give thee the crown of life. %
\vnum{11}He that hath an ear, let him hear what the Spirit saith to the churches. He that overcometh shall not be hurt of the second death.
\subsection*{Letter to Pergamum}
\vnum{12}And to the angel of the church in Pergamum write:

These things saith he that hath the sharp two-edged sword:%
	\cbiblefoot{Isaiah}{49:2}{He hath made my mouth like a sharp sword.} %
\vnum{13}I know where thou dwellest,%
	\cbiblefoot{Ezekiel}{12:2}{Thou dwellest in the midst of the rebellious house, that have eyes to see, and see not, that have ears to hear, and hear not; for they are a rebellious house.}
 even where Satan’s throne is; and thou holdest fast my name, and didst not deny my faith, even in the days of Antipas my witness, my faithful one, who was killed among you, where Satan dwelleth. %
\vnum{14}But I have a few things against thee, because thou hast there some that hold the teaching of Balaam, who taught Balak to cast a stumblingblock before the children of Israel, to eat things sacrificed to idols, and to commit fornication.%
		\cbiblefootduosb{Numbers}{31:15-16}{Moses said unto them, Have ye saved all the women alive? Behold, these caused the children of Israel, through the counsel of Balaam, to commit trespass against Jehovah in the matter of Peor}%
								{25:1-2}{The people began to play the harlot with the daughters of Moab: for they called the people unto the sacrifices of their gods; and the people did eat, and bowed down to their gods.} %
\vnum{15}So hast thou also some that hold the teaching of the Nicolaitans in like manner. %
\vnum{16}Repent therefore; or else I come to thee quickly, and I will make war against them with the sword of my mouth.%
	\cbiblefoot{Jeremiah}{21:5}{I myself will fight against you with an outstretched hand and with a strong arm, even in anger, and in wrath, and in great indignation.} %
\vnum{17}He that hath an ear, let him hear what the Spirit saith to the churches. To him that overcometh, to him will I give of the hidden manna,%
	\cbiblefoot{Exodus}{16:32-34}{Moses said, This is the thing which Jehovah hath commanded, Let an omerful of it be kept throughout your generations, that they may see the bread wherewith I fed you in the wilderness, when I brought you forth from the land of Egypt. And Moses said unto Aaron, Take a pot, and put an omerful of manna therein, and lay it up before Jehovah, to be kept throughout your generations. As Jehovah commanded Moses, so Aaron laid it up before the Testimony, to be kept.} %
and I will give him a white stone, and upon the stone a new name written, which no one knoweth but he that receiveth it.%
	\cbiblefoottriosb{Isaiah}{56:4-5}{Thus saith Jehovah of the eunuchs that keep my sabbaths, and choose the things that please me, and hold fast my covenant: Unto them will I give in my house and within my walls a memorial and a name better than of sons and of daughters; I will give them an everlasting name, that shall not be cut off.}%
					{62:2}{The nations shall see thy righteousness, and all kings thy glory, and thou shalt be called by a new name, which the mouth of Jehovah shall name}
					{65:15}{Ye shall leave your name for a curse unto my chosen; and the Lord Jehovah will slay thee; and he will call his servants by another name.}
\subsection*{Letter to Thyatira}
\vnum{18}And to the angel of the church in Thyatira write:

These things saith the Son of God, who hath his eyes like a flame of fire, and his feet are like unto burnished brass:%
	\footnote{cf. 1:14-15} %
\vnum{19}I know thy works, and thy love and faith and ministry and patience, and that thy last works are more than the first. %
\vnum{20}But I have this against thee, that thou sufferest the woman Jezebel, who calleth herself a prophetess; and she teacheth and seduceth my servants to commit fornication, and to eat things sacrificed to idols.%
	\cbiblefootduo{IKings}{16:31}{As if it had been a light thing for him to walk in the sins of Jeroboam\ldots, he took to wife Jezebel the daughter of Ethbaal king of the Sidonians, and went and served Baal, and worshipped him}%
				 {IIKings}{9:22}{When Joram saw Jehu\ldots he said, Is it peace, Jehu? And he answered, What peace, so long as the whoredoms of thy mother Jezebel and her witchcrafts are so many?} %
\vnum{21}And I gave her time that she should repent; and she willeth not to repent of her fornication. %
\vnum{22}Behold, I cast her into a bed, and them that commit adultery with her into great tribulation, except they repent of her works. %
\vnum{23}And I will kill her children with death; and all the churches shall know that I am he that searcheth the reins and hearts:%
	\cbiblefoot{Jeremiah}{11:20}{O Jehovah of hosts, who judgest righteously, who triest the heart and the mind, I shall see thy vengeance on them.} %
and I will give unto each one of you according to your works. %
\vnum{24}But to you I say, to the rest that are in Thyatira, as many as have not this teaching, who know not the deep things of Satan, as they are wont to say; I cast upon you none other burden. %
\vnum{25}Nevertheless that which ye have, hold fast till I come. %
\vnum{26}And he that overcometh, and he that keepeth my works unto the end, to him will I give authority over the nations: %
\vnum{27}and he shall rule them with a rod of iron, as the vessels of the potter are broken to shivers;%
	\cbiblefoot{Psalms}{2:8-9}{Ask of me, and I will give thee the nations for thine inheritance, and the uttermost parts of the earth for thy possession. Thou shalt break them with a rod of iron; thou shalt dash them in pieces like a potter’s vessel.} %
as I also have received of my Father: %
\vnum{28}and I will give him the morning star. %
\vnum{29}He that hath an ear, let him hear what the Spirit saith~to~the~churches.
