%CHAPTER 14

\chapter{The Lamb and the 144,000 on Mount Zion}
\lettrine[lines=3,slope=0.5em]{\textcolor{red}{A}}{\ nd} I saw%
, and behold, the Lamb standing on the mount Zion,%
	\cbiblefootduo{Joel}{3:16-17}{And Jehovah will roar from Zion, and utter his voice from Jerusalem; and the heavens and the earth shall shake: but Jehovah will be a refuge unto his people, and a stronghold to the children of Israel. So shall ye know that I am Jehovah your God, dwelling in Zion my holy mountain: then shall Jerusalem be holy, and there shall no strangers pass through her any more}%
				  {Isaiah}{4:2-6}{In that day shall the branch of Jehovah be beautiful and glorious, and the fruit of the land shall be excellent and comely for them that are escaped of Israel. And it shall come to pass, that he that is left in Zion, and he that remaineth in Jerusalem, shall be called holy, even every one that is written among the living in Jerusalem; when the Lord shall have washed away the filth of the daughters of Zion, and shall have purged the blood of Jerusalem from the midst thereof, by the spirit of justice, and by the spirit of burning. And Jehovah will create over the whole habitation of mount Zion, and over her assemblies, a cloud and smoke by day, and the shining of a flaming fire by night; for over all the glory shall be spread a covering. And there shall be a pavilion for a shade in the day-time from the heat, and for a refuge and for a covert from storm and from rain}
 and with him a hundred and forty and four thousand, having his name, and the name of his Father, written on their foreheads.%
	\cbiblefoot{Ezekiel}{9:4}{And Jehovah said unto him, Go through the midst of the city, through the midst of Jerusalem, and set a mark upon the foreheads of the men that sigh and that cry over all the abominations that are done in the midst thereof}
\vnum{2} And I heard a voice from heaven, as the voice of many waters, and as the voice of a great thunder: and the voice which I heard was as the voice of harpers harping with their harps: 
\vnum{3} and they sing as it were a new song%
	\footnote{\cbibleref{Psalms}{33:3}{Sing unto him a new song; Play skilfully with a loud noise}\ldots %
			  \cbiblechvs{Psalms}{40:3}{And he hath put a new song in my mouth, even praise unto our God: Many shall see it, and fear, And shall trust in Jehovah}\ldots %
			  \cbiblechvs{Psalms}{96:1}{Oh sing unto Jehovah a new song: Sing unto Jehovah, all the earth}\ldots %
			  \cbiblechvs{Psalms}{98:1}{Oh sing unto Jehovah a new song; For he hath done marvellous things: His right hand, and his holy arm, hath wrought salvation for him}\ldots %
			  \cbiblechvs{Psalms}{144:9}{I will sing a new song unto thee, O God: Upon a psaltery of ten strings will I sing praises unto thee}\ldots %
			  \cbiblechvs{Psalms}{149:1}{Praise ye Jehovah. Sing unto Jehovah a new song, And his praise in the assembly of the saints}\ldots %
			  \cbibleref{Isaiah}{42:10}{Sing unto Jehovah a new song, and his praise from the end of the earth; ye that go down to the sea, and all that is therein, the isles, and the inhabitants thereof}\ldots}
 before the throne, and before the four living creatures and the elders: and no man could learn the song save the hundred and forty and four thousand, even they that had been purchased out of the earth. 
\vnum{4} These are they that were not defiled with women; for they are virgins. These are they that follow the Lamb whithersoever he goeth. These were purchased from among men, to be the first-fruits unto God and unto the Lamb. 
\vnum{5} And in their mouth was found no lie:%
	\cbiblefoottrio{Zephaniah}{3:13}{The remnant of Israel shall not do iniquity, nor speak lies; neither shall a deceitful tongue be found in their mouth; for they shall feed and lie down, and none shall make them afraid}%
				   {Isaiah}{53:9}{And they made his grave with the wicked, and with a rich man in his death; although he had done no violence, neither was any deceit in his mouth}%
				   {Psalms}{32:2}{Blessed is the man unto whom Jehovah imputeth not iniquity,
And in whose spirit there is no guile}
 they are without blemish.
\subsection*{The Three Angels}
\vnum{6} And I saw another angel flying in mid heaven, having eternal good tidings to proclaim unto them that dwell on the earth, and unto every nation and tribe and tongue and people; 
\vnum{7} and he saith with a great voice, Fear God, and give him glory; for the hour of his judgment is come: and worship him that made the heaven and the earth and sea and fountains of waters.

\vnum{8} And another, a second angel, followed, saying, Fallen, fallen is Babylon the great,%
	\cbiblefootduo{Isaiah}{21:9}{And he answered and said, Fallen, fallen is Babylon; and all the graven images of her gods are broken unto the ground}%
				  {Jeremiah}{51:6-9}{Flee out of the midst of Babylon, and save every man his life; be not cut off in her iniquity: for it is the time of Jehovah’s vengeance; he will render unto her a recompense. Babylon hath been a golden cup in Jehovah’s hand, that made all the earth drunken: the nations have drunk of her wine; therefore the nations are mad. Babylon is suddenly fallen and destroyed: wail for her; take balm for her pain, if so be she may be healed. We would have healed Babylon, but she is not healed: forsake her, and let us go every one into his own country; for her judgment reacheth unto heaven, and is lifted up even to the skies}
 that hath made all the nations to drink of the wine of the wrath of her fornication.

\vnum{9} And another angel, a third, followed them, saying with a great voice, If any man worshippeth the beast and his image, and receiveth a mark on his forehead, or upon his hand, 
\vnum{10} he also shall drink of the wine of the wrath of God, which is prepared unmixed in the cup of his anger;%
	\cbiblefoottrio{Isaiah}{51:17-23}{Awake, awake, stand up, O Jerusalem, that hast drunk at the hand of Jehovah the cup of his wrath; thou hast drunken the bowl of the cup of staggering, and drained it\ldots Therefore hear now this, thou afflicted, and drunken, but not with wine: Thus saith thy Lord Jehovah, and thy God that pleadeth the cause of his people, Behold, I have taken out of thy hand the cup of staggering, even the bowl of the cup of my wrath; thou shalt no more drink it again: and I will put it into the hand of them that afflict thee\ldots}%
				   {Jeremiah}{25:15-29}{For thus saith Jehovah, the God of Israel, unto me: Take this cup of the wine of wrath at my hand, and cause all the nations, to whom I send thee, to drink it. And they shall drink, and reel to and fro, and be mad, because of the sword that I will send among them. Then took I the cup at Jehovah’s hand, and made all the nations to drink, unto whom Jehovah had sent me\ldots to make them a desolation, an astonishment, a hissing, and a curse, as it is this day\ldots Drink ye, and be drunken, and spew, and fall, and rise no more, because of the sword which I will send among you. And it shall be, if they refuse to take the cup at thy hand to drink, then shalt thou say unto them, Thus saith Jehovah of hosts: Ye shall surely drink. For, lo, I begin to work evil at the city which is called by my name; and should ye be utterly unpunished? Ye shall not be unpunished; for I will call for a sword upon all the inhabitants of the earth, saith Jehovah of hosts}%
				   {Psalms}{75:8}{For in the hand of Jehovah there is a cup, and the wine foameth; It is full of mixture, and he poureth out of the same: Surely the dregs thereof, all the wicked of the earth shall drain them, and drink them}
 and he shall be tormented with fire and brimstone%
	\cbiblefoottrio{Genesis}{19:24}{Then Jehovah rained upon Sodom and upon Gomorrah brimstone and fire from Jehovah out of heaven}%
					{Psalms}{11:6}{Upon the wicked he will rain snares; Fire and brimstone and burning wind shall be the portion of their cup}%
					{Ezekiel}{38:22}{And with pestilence and with blood will I enter into judgment with him; and I will rain upon him, and upon his hordes, and upon the many peoples that are with him, an overflowing shower, and great hailstones, fire, and brimstone}%
 in the presence of the holy angels, and in the presence of the Lamb: 
\vnum{11} and the smoke of their torment goeth up for ever and ever;%
	\cbiblefoot{Isaiah}{34:8-10}{For Jehovah hath a day of vengeance, a year of recompense for the cause of Zion. And the streams of Edom shall be turned into pitch, and the dust thereof into brimstone, and the land thereof shall become burning pitch. It shall not be quenched night nor day; the smoke thereof shall go up for ever; from generation to generation it shall lie waste; none shall pass through it for ever and ever}
 and they have no rest day and night,%
	\cbiblefoot{Isaiah}{48:22}{There is no peace, saith Jehovah, to the wicked}
 they that worship the beast and his image, and whoso receiveth the mark of his name. 
\vnum{12} Here is the patience of the saints, they that keep the commandments of God, and the faith of Jesus.

\vnum{13} And I heard a voice from heaven saying, Write, Blessed are the dead who die in the Lord from henceforth: yea, saith the Spirit, that they may rest from their labors; for their works follow with them.
\subsection*{Harvesting the Earth and Trampling the Winepress}
\vnum{14} And I saw, and behold, a white cloud; and on the cloud I saw one sitting like unto a son of man,%
	\cbiblefoot{Daniel}{7:13-14}{ I saw in the night-visions, and, behold, there came with the clouds of heaven one like unto a son of man, and he came even to the ancient of days, and they brought him near before him. And there was given him dominion, and glory, and a kingdom, that all the peoples, nations, and languages should serve him: his dominion is an everlasting dominion, which shall not pass away, and his kingdom that which shall not be destroyed}
 having on his head a golden crown, and in his hand a sharp sickle. 
\vnum{15} And another angel came out from the temple, crying with a great voice to him that sat on the cloud, Send forth thy sickle, and reap: for the hour to reap is come; for the harvest of the earth is ripe.%
	\cbiblefootduo{Joel}{3:13}{Put ye in the sickle; for the harvest is ripe: come, tread ye; for the winepress is full, the vats overflow; for their wickedness is great}%
				  {Jeremiah}{51:33}{For thus saith Jehovah of hosts, the God of Israel: The daughter of Babylon is like a threshing-floor at the time when it is trodden; yet a little while, and the time of harvest shall come for her}
\vnum{16} And he that sat on the cloud cast his sickle upon the earth; and the earth was reaped.

\vnum{17} And another angel came out from the temple which is in heaven, he also having a sharp sickle. 
\vnum{18} And another angel came out from the altar, he that hath power over fire; and he called with a great voice to him that had the sharp sickle, saying, Send forth thy sharp sickle, and gather the clusters of the vine of the earth; for her grapes are fully ripe. 
\vnum{19} And the angel cast his sickle into the earth, and gathered the vintage of the earth, and cast it into the winepress, the great winepress, of the wrath of God. 
\vnum{20} And the winepress was trodden without the city, and there came out blood from the winepress,%
	\cbiblefootduo{Isaiah}{63:1-6}{Who is this that cometh from Edom, with dyed garments from Bozrah? This that is glorious in his apparel, marching in the greatness of his strength? I that speak in righteousness, mighty to save. Wherefore art thou red in thine apparel, and thy garments like him that treadeth in the winevat? I have trodden the winepress alone; and of the peoples there was no man with me: yea, I trod them in mine anger, and trampled them in my wrath; and their lifeblood is sprinkled upon my garments, and I have stained all my raiment. For the day of vengeance was in my heart, and the year of my redeemed is come. And I looked, and there was none to help; and I wondered that there was none to uphold: therefore mine own arm brought salvation unto me; and my wrath, it upheld me. And I trod down the peoples in mine anger, and made them drunk in my wrath, and I poured out their lifeblood on the earth.}%
				{Jeremiah}{25:30-31}{Therefore prophesy thou against them all these words, and say unto them, Jehovah will roar from on high, and utter his voice from his holy habitation; he will mightily roar against his fold; he will give a shout, as they that tread the grapes, against all the inhabitants of the earth. A noise shall come even to the end of the earth; for Jehovah hath a controversy with the nations; he will enter into judgment with all flesh: as for the wicked, he will give them to the sword, saith Jehovah}
 even unto the bridles of the horses, as far as a thousand and six hundred furlongs.