%CHAPTER 22
%\chapterfigure[scale=0.2]{Durer/Durer_Revelation_Four_Riders.jpg}{The Four Horsemen. Albrecht Dürer, 1498.}

\chapter{The River and the Tree of Life}
\lettrine[lines=3,slope=0.5em]{\textcolor{red}{A}}{\ nd} he showed me a river of water%
	\footnote{%
		\cbibleref{Genesis}{2:10}{A river went out of Eden to water the garden; and from thence it was parted, and became four heads} %
		\cbibleref{Ezekiel}{47:1}{He brought me back unto the door of the house; and, behold, waters issued out from under the threshold of the house eastward} %
		\cbibleref{Zechariah}{14:8}{It shall come to pass in that day, that living waters shall go out from Jerusalem; half of them toward the eastern sea, and half of them toward the western sea: in summer and in winter shall it be} %
		\cbibleref{Psalms}{46:4}{There is a river, the streams whereof make glad the city of God, the holy place of the tabernacles of the Most High.}%
	} %
of life, bright as crystal, proceeding out of the throne of God and of the Lamb, 
\vnum{2}in the midst of the street thereof. And on this side of the river and on that%
	\cbiblefoot{Ezekiel}{47:12}{By the river upon the bank thereof, on this side and on that side, shall grow every tree for food, whose leaf shall not wither, neither shall the fruit thereof fail: it shall bring forth new fruit every month, because the waters thereof issue out of the sanctuary; and the fruit thereof shall be for food, and the leaf thereof for healing.} %
was the tree of life,%
	\cbiblefoot{Genesis}{2:9}{Out of the ground made Jehovah God to grow every tree that is pleasant to the sight, and good for food; the tree of life also in the midst of the garden.} %
bearing twelve manner of fruits, yielding its fruit every month: and the leaves of the tree were for the healing of the nations.%	
	\cbiblefoot{Jeremiah}{3:17}{At that time they shall call Jerusalem the throne of Jehovah; and all the nations shall be gathered unto it, to the name of Jehovah, to Jerusalem: neither shall they walk any more after the stubbornness of their evil heart.} %
\vnum{3}And there shall be no curse any more:%
	\cbiblefoot{Zechariah}{14:11}{Men shall dwell therein, and there shall be no more curse; but Jerusalem shall dwell safely.} %
and the throne of God and of the Lamb shall be therein: and his servants shall serve him; 
\vnum{4}and they shall see his face;%
	\cbiblefoot{Psalms}{17:15}{As for me, I shall behold thy face in righteousness; I shall be satisfied, when I awake, with beholding thy form.} %
and his name shall be on their foreheads. 
\vnum{5}And there shall be night no more; and they need no light of lamp, neither light of sun; for the Lord God shall give them light:%
	\cbiblefoot{Isaiah}{60:19}{The sun shall be no more thy light by day; neither for brightness shall the moon give light unto thee: but Jehovah will be unto thee an everlasting light, and thy God thy glory.} %
and they shall reign for ever and ever.%
	\cbiblefootduosbch{Daniel}{7:18}{The saints of the Most High shall receive the kingdom, and possess the kingdom for ever, even for ever and ever}%
					{7:27}{The kingdom and the dominion, and the greatness of the kingdoms under the whole heaven, shall be given to the people of the saints of the Most High: his kingdom is an everlasting kingdom, and all dominions shall serve and obey him.} %

\vnum{6}And he said unto me, These words are faithful and true: and the Lord, the God of the spirits of the prophets, sent his angel to show unto his servants the things which must shortly come to pass.%
	\cbiblefootduosbch{Daniel}{2:28-29}{But there is a God in heaven that revealeth secrets, and he hath made known to the king Nebuchadnezzar what shall be in the latter days. Thy dream, and the visions of thy head upon thy bed, are these: as for thee, O king, thy thoughts came into thy mind upon thy bed, what should come to pass hereafter; and he that revealeth secrets hath made known to thee what shall come to pass}%
			{2:45}{Forasmuch as thou sawest that a stone was cut out of the mountain without hands, and that it brake in pieces the iron, the brass, the clay, the silver, and the gold; the great God hath made known to the king what shall come to pass hereafter: and the dream is certain, and the interpretation thereof sure.} %
\vnum{7}And behold, I come quickly.%
	\cbiblefoot{Isaiah}{40:10}{Behold, the Lord Jehovah will come as a mighty one, and his arm will rule for him: Behold, his reward is with him, and his recompense before him.} %
Blessed is he that keepeth the words of the prophecy of this book.

\vnum{8}And I John am he that heard and saw these things. And when I heard and saw, I fell down to worship before the feet of the angel that showed me these things. 
\vnum{9}And he saith unto me, See thou do it not: I am a fellow-servant with thee and with thy brethren the prophets, and with them that keep the words of this book: worship God.

\vnum{10}And he saith unto me, Seal not up the words of the prophecy of this book; for the time is at hand. 
\vnum{11}He that is unrighteous, let him do unrighteousness still: and he that is filthy, let him be made filthy still:%
	\cbiblefootduo{Ezekiel}{3:27}{When I speak with thee, I will open thy mouth, and thou shalt say unto them, Thus saith the Lord Jehovah: He that heareth, let him hear; and he that forbeareth, let him forbear: for they are a rebellious house}%
					{Daniel}{12:10}{Many shall purify themselves, and make themselves white, and be refined; but the wicked shall do wickedly; and none of the wicked shall understand; but they that are wise shall understand.} %
and he that is righteous, let him do righteousness still: and he that is holy, let him be made holy still.%
	\cbiblefoot{Isaiah}{56:1}{Thus saith Jehovah, Keep ye justice, and do righteousness; for my salvation is near to come, and my righteousness to be revealed.} %
\vnum{12}Behold, I come quickly; and my reward is with me, to render to each man according as his work is. 
\vnum{13}I am the Alpha and the Omega, the first and the last, the beginning and the end. 
\vnum{14}Blessed are they that wash their robes, that they may have the right to come to the tree of life, and may enter in by the gates into the city.%
	\cbiblefoot{Psalms}{118:19-20}{Open to me the gates of righteousness: I will enter into them, I will give thanks unto Jehovah. This is the gate of Jehovah; the righteous shall enter into it.} %
\vnum{15} Without are the dogs, and the sorcerers, and the fornicators, and the murderers, and the idolaters, and every one that loveth and maketh a lie.

\vnum{16}I Jesus have sent mine angel%
	\cbiblefoot{Malachi}{3:1}{Behold, I send my messenger, and he shall prepare the way before me: and the Lord, whom ye seek, will suddenly come to his temple; and the messenger of the covenant, whom ye desire, behold, he cometh, saith Jehovah of hosts.} %
to testify unto you these things for the churches. I am the root and the offspring of David, the bright, the morning star.%
	\cbiblefoot{Numbers}{24:17}{I see him, but not now; I behold him, but not nigh: there shall come forth a star out of Jacob, and a sceptre shall rise out of Israel, and shall smite through the corners of Moab, and break down all the sons of tumult.} %

\vnum{17}And the Spirit and the bride say, Come. And he that heareth, let him say, Come. And he that is athirst, let him come: he that will, let him take the water of life freely.%
	\cbiblefoot{Isaiah}{55:1}{Ho, every one that thirsteth, come ye to the waters, and he that hath no money; come ye, buy, and eat; yea, come, buy wine and milk without money and without price.} 

\vnum{18}I testify unto every man that heareth the words of the prophecy of this book, If any man shall add unto them,%
	\cbiblefoot{Deuteronomy}{4:2}{Ye shall not add unto the word which I command you, neither shall ye diminish from it, that ye may keep the commandments of Jehovah your God which I command you.} %
God shall add unto him the plagues which are written in this book:%
	\cbiblefoot{Deuteronomy}{29:20-21}{Jehovah will not pardon him, but then the anger of Jehovah and his jealousy will smoke against that man, and all the curse that is written in this book shall lie upon him, and Jehovah will blot out his name from under heaven. And Jehovah will set him apart unto evil out of all the tribes of Israel, according to all the curses of the covenant that is written in this book of the law.} %
\vnum{19}and if any man shall take away from the words of the book of this prophecy, God shall take away his part from the tree of life, and out of the holy city, which are written in this book.

\vnum{20}He who testifieth these things saith, Yea: I come quickly. Amen: come, Lord Jesus.

\vnum{21}The grace of the Lord Jesus be with the saints. Amen.
