%CHAPTER 5
\chapter{The Scroll and the Lamb}
\lettrine[lines=3,slope=0.5em]{\textcolor{red}{A}}{\ nd} I saw in the right hand of him that sat on the throne a book written within and on the back, close sealed with seven seals.%
 \cbiblefoottrio{Ezekiel}{2:9-10}{A scroll of a book was therein; and he spread it before me: and it was written within and without; and there were written therein lamentations, and mourning, and woe}%
 					{Isaiah}{29:11-12}{All vision is become unto you as the words of a book that is sealed, which men deliver to one that is learned, saying, Read this, I pray thee; and he saith, I cannot, for it is sealed: and the book is delivered to him that is not learned, saying, Read this, I pray thee; and he saith, I am not learned}%
 					{Daniel}{12:4, 9}{But thou, O Daniel, shut up the words, and seal the book, even to the time of the end\ldots Go thy way, Daniel; for the words are shut up and sealed till the time of the end.} %
\vnum{2}And I saw a strong angel proclaiming with a great voice, Who is worthy to open the book, and to loose the seals thereof? %
\vnum{3}And no one in the heaven, or on the earth, or under the earth, was able to open the book, or to look thereon. %
\vnum{4}And I wept much, because no one was found worthy to open the book, or to look thereon: %
\vnum{5}and one of the elders saith unto me, Weep not; behold, the Lion that is of the tribe of Judah,%
	\cbiblefoot{Genesis}{49:9}{Judah is a lion's whelp; from the prey, my son, thou art gone up: He stooped down, he couched as a lion, and as a lioness; who shall rouse him up?}
 the Root of David%
 	\cbiblefoot{Isaiah}{11:1, 10}{And there shall come forth a shoot out of the stock of Jesse, and a branch out of his roots shall bear fruit\ldots And it shall come to pass in that day, that the root of Jesse, that standeth for an ensign of the peoples, unto him shall the nations seek; and his resting-place shall be glorious.}
 , hath overcome to open the book and the seven seals thereof. %
\vnum{6}And I saw in the midst of the throne and of the four living creatures, and in the midst of the elders, a Lamb standing, as though it had been slain,%
	\cbiblefootduo{Isaiah}{53:7}{As a lamb that is led to the slaughter, and as a sheep that before its shearers is dumb, so he opened not his mouth}%
					  {Jeremiah}{11:19}{I was like a gentle lamb that is led to the slaughter}
 having seven horns, and seven eyes, which are the seven Spirits of God, sent forth into all the earth.%
 	\cbiblefoot{Zechariah}{4:10}{These seven shall rejoice\ldots these are the eyes of Jehovah, which run to and fro through the whole earth.} %
\vnum{7}And he came, and he taketh it out of the right hand of him that sat on the throne. %
\vnum{8}And when he had taken the book, the four living creatures and the four and twenty elders fell down before the Lamb, having each one a harp, and golden bowls full of incense, which are the prayers of the saints.%
	\cbiblefoot{Psalms}{141:2}{Let my prayer be set forth as incense before thee; the lifting up of my hands as the evening sacrifice.} %
\vnum{9}And they sing a new song,%
	\footnote{\cbibleref{Psalms}{33:3}{Sing unto him a new song; Play skilfully with a loud noise}\ldots%
			\cbiblechvs{Psalms}{40:3}{And he hath put a new song in my mouth, even praise unto our God: many shall see it, and fear, And shall trust in Jehovah}\ldots%
			\cbiblechvs{Psalms}{96:1}{Oh sing unto Jehovah a new song: sing unto Jehovah, all the earth}\ldots%
			\cbiblechvs{Psalms}{98:1}{Oh sing unto Jehovah a new song; for he hath done marvellous things: His right hand, and his holy arm, hath wrought salvation for him}\ldots%
			\cbiblechvs{Psalms}{144:9}{I will sing a new song unto thee, O God: upon a psaltery of ten strings will I sing praises unto thee.}\ldots%
			\cbiblechvs{Psalms}{149:1}{Praise ye Jehovah. Sing unto Jehovah a new song, and his praise in the assembly of the saints}\ldots%
			\cbibleref{Isaiah}{42:10}{Sing unto Jehovah a new song, and his praise from the end of the earth; ye that go down to the sea, and all that is therein, the isles, and the inhabitants thereof.}
	}
 saying,
\begin{verse}
Worthy art thou to take the book,\\*
\vin and to open the seals thereof:\\* 
for thou wast slain,\\*
\vin and didst purchase unto God with thy blood\\* 
\vin men of every tribe, and tongue, and people, and nation,\\* %
\vleftofline{\vnum{10}}and madest them to be unto our God a kingdom and priests;%
	\cbiblefootduo{Exodus}{19:5-6}{Ye shall be mine own possession from among all peoples\ldots and ye shall be unto me a kingdom of priests, and a holy nation.}%
				{Isaiah}{61:6-7}{Ye shall be named the priests of Jehovah; men shall call you the ministers of our God: ye shall eat the wealth of the nations, and in their glory shall ye boast yourselves. Instead of your shame ye shall have double; and instead of dishonor they shall rejoice in their portion: therefore in their land they shall possess double; everlasting joy shall be unto them.}\\*
\vin and they reign upon the earth.
\end{verse}

\vnum{11}And I saw, and I heard a voice of many angels round about the throne and the living creatures and the elders; and the number of them was ten thousand times ten thousand, and thousands of thousands;%
	\cbiblefoot{Daniel}{7:9-10}{I beheld till thrones were placed, and one that was ancient of days did sit\ldots Thousands of thousands ministered unto him, and ten thousand times ten thousand stood before him: the judgment was set, and the books were opened.}
\vnum{12}saying with a great voice,
\begin{verse}
Worthy is the Lamb that hath been slain to receive\\*\vin the power, and riches, and wisdom,\\*\vin and might, and honor, and glory, and blessing.
\end{verse}
\vnum{13}And every created thing which is in the heaven, and on the earth, and under the earth, and on the sea, and all things that are in them, heard I saying,
\begin{verse}
Unto him that sitteth on the throne, and unto the Lamb,\\*\vin be the blessing, and the honor, and the glory, and the dominion,\\*\vin for ever and ever.
\end{verse}
\vnum{14}And the four living creatures said, Amen. And the elders fell down and worshipped.
